# Week 1 {-}

## Welcome to the New Season {-}

Well we're back with an exciting start to the new season. Manchester United turned in their best performance in  years and Huddersfield had a dream return to the top tier. Some things remain the same though with Dele Alli conning Shelvey into a dismissal and Liverpool conceding from corners

### Game of the Week {-}

Arsenal and Leicester swapping blows with the Gunners emerging victorious

```{r topGame_01}

embed_youtube(
  "tXq8TlZupEY",
  width = 420,
  height = 315,
  frameborder = 0,
  allowfullscreen = TRUE,
  query = NULL
  ) %>%
  div(align = "center")
  

```






## English born Players {-}

One fairly constant theme has been the downturn in English player over the years. 



```{r english_starters_01}

games <- teamGames %>%
  ungroup() %>%
  filter(tmYrGameOrder == 1) %>%
  pull(MATCHID)

playerGame %>%
  filter(MATCHID %in% games & START > 0) %>%
  group_by(season, COUNTRY) %>%
  summarise(tot = n()) %>%
  group_by(season) %>%
  mutate(pc = round(100 * tot / sum(tot), 1)) %>%
  filter(COUNTRY == "England") %>%
  ungroup() %>%
  plot_ly(x =  ~ season, y =  ~ pc) %>%
  add_lines() %>%
  layout(
    margin = list(b = 70),
    title = "% of Starters born in England - EPL Round 1",
    xaxis = list(title = ""),
    yaxis = list(title = "% ")
  ) %>%  config(displayModeBar = F, showLink = F)

```

This appears to have stabilized over the past couple of seasons

---

Check out how each of the teams started the season


```{r country_table_01}


# 1st weeek this year
playerGame %>%
  filter(MATCHID %in% games & START > 0 & season == "2017/18") %>%
  mutate(place = paste0(city, ", ", COUNTRY)) %>%
  select(team = TEAMNAME, player = LASTNAME, place) %>%
  arrange(team, player) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE,
      lengthMenu = c(11, 11)
    )
  )


```

New-signing, Lacazette, was the only Frenchman in Arsenal's starting line-up (It has been as high as seven). 

 Stoke boasted ten different nationalities
 
---

Here is the breakdown across the EPL

```{r allcountries_01, fig.height=10}

country.df <- playerGame %>%
  filter(MATCHID %in% games & START > 0 & season == "2017/18") %>%
  group_by(COUNTRY) %>%
  tally()

#use forcats - works even though character - then prder on reverse alphabetically - ideally opposite of that
country.df  %>%
  plot_ly(x =  ~ n,
          y =  ~ fct_reorder(COUNTRY, n),
          height = 1000) %>%
  layout(
    margin = list(l = 100, pad = 5),
    yaxis = list(title = ""),
    xaxis = list(title = "")
  ) %>%
  config(displayModeBar = F, showLink = F)


```


No clear candidate for second place but newly-promoted Huddersfield coach, David Wagner, has signed a few fellow-Germans

Only two Scots, Fraser (Bournemouth) and 33-year-old, Fletcher (Stoke). The last time Scotland held the second rank, after England, in a game to lead off a season was in 1999





## Big Away Wins {-}


<p class="factoid">What Birmingham never did in seven seasons and Southampton only achieved at the 186th attempt, Huddersfield accomplished at the first time of asking<br>Won their first EPL away game by at least a three goal margin<br>Surprisingly, three other teams match them, including Blackpool, who were relegated the same year</p> 



```{r big_win_01, fig.height= 12}



# first games for teams that have had 3 goals margins
didIt <- standings %>%
  arrange(gameDate) %>%
  filter(venue == "A") %>%
  group_by(team) %>%
  mutate(awayGameOrder = row_number()) %>%
  filter((GF - GA) >= 3) %>%
  group_by(team) %>%
  slice(1)

# teams that have never achieved
never <- setdiff(unique(standings$team), unique(didIt$team))

# never teams away games to date - select relevant cariables+ add ever
neverTeams <- standings %>%
  filter(team %in% never & venue == "A") %>%
  group_by(team) %>%
  arrange(tmGameOrder) %>%
  mutate(awayGameOrder = row_number()) %>%
  arrange(desc(awayGameOrder)) %>%
  slice(1) %>%
  select(team, awayGameOrder) %>%
  mutate(ever = "Never")

# cf above re fields
didItTeams <- didIt %>%
  select(team, awayGameOrder) %>%
  mutate(ever = "First")

# combine
df <- rbind(neverTeams, didItTeams)

# order team as factor for chart
df$team <-
  factor(df$team, levels = df$team[order(df$awayGameOrder)])


#chart
df %>%
  plot_ly(x =  ~ awayGameOrder,
          y =  ~ team,
          height = 1000) %>%
  add_bars(color =  ~ ever) %>%
  layout(
    title = " EPL Away Games before 3 goal victory",
    margin = list(l = 120, pad = 5),
    xaxis = list(title = "Games"),
    yaxis = list(title = "")
    
  ) %>%  config(displayModeBar = F, showLink = F)

```

It took Newcastle more than a century of matches which is a touch unexpected given that they were a high-scoring, successful team when they first made the Premier league

Southampton definitely buck the trend taking 186 games to reach the target - though they have had more success recently

```{r southampton_01}


standings %>%
  filter(venue == "A" & team == "Southampton") %>%
  mutate(GD = GF - GA) %>%
  group_by(GD) %>%
  tally() %>%
  plot_ly(x =  ~ GD, y =  ~ n) %>%
  add_bars(color = I("red3")) %>%
  layout(
    title = "Southampton PL results by Goal Difference",
    xaxis = list(title = "Goal Difference"),
    yaxis = list(title = "Games Count")
  ) %>%  config(displayModeBar = F, showLink = F)


```

## Man Utd are back {-}

I know, only West Ham, but still far from the turgid play of recent seasons. They end by far their longest run without a four goal victory


```{r mnu_streak_01}


standings %>%
  arrange(gameDate) %>%
  group_by(team) %>%
  filter((GF - GA) >= 4 & team == "Man. Utd.") %>%
  select(gameDate, tmGameOrder) %>%
  mutate(last = lag(tmGameOrder)) %>%
  mutate(last = (ifelse(is.na(last), 0, last)),
         diff = tmGameOrder - last,
         order = row_number() - 1) %>%
  plot_ly(x =  ~ order, y =  ~ diff) %>%
  add_bars(color = I("red")) %>%
  layout(
    title = "Man Utd sequences between 4 goal victories",
    xaxis = list(title = ""),
    yaxis = list(title = "")
  ) %>%  config(displayModeBar = F, showLink = F)

```

---

### Totally Football and Guardian Weekly Review Podcasts {-}




<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6203313-chelsea-crash-and-burn-ley-as-huddersfield-make-their-mark/embed/v4?eid=AQAAAO9tlFmxp14A" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6203084-goals-petulance-and-absent-defences-the-premier-league-is-back-football-weekly/embed/v4?eid=AQAAANRslFnMpl4A" title="audioBoom player"></iframe>

---

### Results and Table {-}

```{r results_01}

hth %>%
  filter(season == "2017/18" &
           gameDate >= "2017-08-09" &
           gameDate <= "2017-08-15") %>% 
  filter(venue == "H") %>%
  arrange(team) %>%
  select(Home = team, GF, GA, Away = OppTeam) %>%
  DT::datatable(
    rownames = FALSE,
    class = 'compact stripe hover row-border',
    colnames = c('', '', '', ''),
    options = list(
      paging = FALSE,
      searching = FALSE,
      info = FALSE,
      columnDefs = list(list(
        width = '40%', targets = list(0, 3)
      ))
    ),
    width = 250
  )
```

---

```{r table_01}



# table
hth %>%
  filter(season == "2017/18" & gameDate <= "2017-08-17") %>%
  group_by(team) %>%
  mutate(
    W = ifelse(res == "Win", 1, 0),
    L = ifelse(res == "Loss", 1, 0),
    D = ifelse(res == "Draw", 1, 0)
  ) %>%
  summarise(
    P = n(),
    Pts = sum(points),
    W = sum(W),
    D = sum(D),
    L = sum(L),
    GD = sum(GF) - sum(GA),
    GF = sum(GF)
  ) %>%
  arrange(desc(Pts), desc(GD), desc(GF), team) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    colnames = c('', 'P', 'Pts', 'W', 'D', 'L', 'GD', 'GF'),
    rownames = TRUE,
    options = list(
      paging = FALSE,
      searching = FALSE,
      info = FALSE
    )
  )


```




### Final Factoid {-}

<p class="factoid">Crystal Palace have won just nine of their most recent 33 league games. More than half these victories came in  the 7 games in which Sakho played a full ninety minutes</p> 

***

## End of Season Update {-}

### English born Players {-}

No great departure from week 1, with a slight uptick for the second year running

```{r byYear_02u}
games <- teamGames %>%
  ungroup() %>%
  #filter(tmYrGameOrder == 1) %>%
  pull(MATCHID)

playerGame %>%
  filter(MATCHID %in% games & START > 0) %>%
  group_by(season, COUNTRY) %>%
  summarise(tot = n()) %>%
  group_by(season) %>%
  mutate(pc = round(100 * tot / sum(tot), 1)) %>%
  filter(COUNTRY == "England") %>%
  ungroup() %>%
  plot_ly(x =  ~ season, y =  ~ pc) %>%
  add_lines() %>%
  layout(
    margin = list(b = 70),
    title = "% of Starters born in England  by Season",
    xaxis = list(title = ""),
    yaxis = list(title = "% ")
  ) %>%  config(displayModeBar = F, showLink = F)
```

There is a tremendous range, by club. Here is a chart ordering clubs by finishing position

```{r by_team_position_02u, fig.height = 8}
games <- teamGames %>%
  ungroup() %>%
  filter(season=="2017/18") %>%
  pull(MATCHID)

finalPos <- standings %>% 
  filter(season=="2017/18",tmYrGameOrder==1) %>% 
  select(team,final_Pos)



test <- playerGame %>%
  filter(MATCHID %in% games & START > 0, COUNTRY == "England") %>%
  group_by(TEAMNAME, COUNTRY) %>%
  summarise(pc = round(100*n()/418,1)) %>%
  left_join(finalPos,by=c("TEAMNAME"="team")) 



test %>% 
  plot_ly(x= ~ pc,y=~fct_reorder(TEAMNAME, -final_Pos)
          ) %>% 
  add_markers() %>% 
  layout(paper_bgcolor='transparent',plot_bgcolor='transparent',
         margin=list(l=120),
         title="Proportion of English-born Starters by Club - 2017/18",
         yaxis=list(title=""),
         xaxis=list(title= "% English Born")) %>% config(displayModeBar = F, showLink = F)
  



```

 Unsurprisingly, the top teams tend to be the wealthiest, who can afford to bring in the best players from around the world

Chelseas 6.2% does make the top 5 in terms of proprtion of non-English for any season but the record is Arsenal's 2007/8 team with only 3.1%. A teenage, Theo Walcott was the only Englishman to feature for more than the 155 minutes of Justin Hoyte

The teams relegated are fairly typical so any shift may depend on the promoted club and Wolves have a lot of Portugese players 


***

### Man Utd are back {-}

United really turned it on early in the season and immediately followed up with another 4-0 win at Swansea in the next fixture, with equivalent scorelines against Everton and Crystal Palace by week 7. However, that was that and they go into next year on a 31 game drought (equal 6th worst in their PL history)

The team scored 5 or more on 44 occasions during Sir Alex Ferguson's 21 year reign in the Premier League. They have not achieved it once since

***

## Code Tips {-}

 * dplyr - pull, slice
 * forcats - fct_reorder
 * DT - columnDefs, lengthMenu