# Week 8 {-}

## Game of the week Crystal Palace 2 - 1 Chelsea {-}

Champions lose to previously goalless, pointless London rivals


```{r topGame_08}

embed_youtube("MuLTnBYuEWs", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```


<p class="factoid">Julian Speroni and Gareth McAuley made their first appearances of the season becoming the oldest player and oldest outfielder, respectively, to appear in this campaign</p>



***

## Mo's MO {-}

It is a well known that Jose Mourinho's modus operandi in away games against top sides is to play very conseravtively. It was not that succesful last season and although he got a point at Liverpool, the latter's poor defence was probably there for the taking against a team that had averaged three goals per game after seven rounds

Jamie Redknapp mentioned that you don't expect many 3-2 victories from a Mourinho team.     

 In fact, there have been six

```{r }
shortcode("tweet", "919708190715002881")
```

Well behind Sir Alex Fergsuon against whom all managers are measured and who has a marginally better overall win percentage (53-51)

Let's look at a heatmaps of their results. Hover squares for totals

```{r no_score_08}

temp <- managerGame %>%
  filter(ManagerID == "MourinhoJ") %>%
  mutate(combo = paste0(GF, GA)) %>%
  group_by(combo) %>%
  tally()


allCombos <- expand.grid(data.frame(GF = 0:9, GA = 0:9)) %>%
  mutate(combo = paste0(GF, GA)) #still a df with 100vals

test <- allCombos %>%
  left_join(temp) %>%
  select(GF, GA, count = n)

# need to transform
Games <- t(matrix(
  test$count,
  nrow = 10,
  ncol = 10,
  byrow = TRUE,
  dimnames = list(unique(test$GF),
                  unique(test$GA))
))


plot_ly(
  x = unique(test$GF),
  y = unique(test$GF),
  z = Games,
  key = Games,
  hoverinfo = "z",
  colorscale = 'YIOrRd',
  reversescale = T,
  type = "heatmap"
) %>%
  layout(
    title = "Mourinho Results by Goal Score",
    xaxis = list(title = "Goals Against"),
    yaxis = list(title = "Goals For")
  ) %>%  config(displayModeBar = F, showLink = F)

```


So his 23rd goalless stalemate. 2-0 victories (40) is narrowly the most common outcome


```{r no_score_AF_08}

```

Interestingly, 1-0 victories is the only result that topped the century mark but 0-0 draws were relatively less commonplace than with Mourinho 

***

## City on the Rampage {-}

Meanwhile, United's cross-city rivals continue apace with only their second-ever seven goal haul in the Premier League (7-0 v Norwich 2103/4) and now boast the most ever goals scored after eight rounds

```{r cum_GF_08}



prev <- standings %>%
  filter(tmYrGameOrder <= 8 & season < "2017/18") %>%
  select(team, tmYrGameOrder, season, cumGF) %>%
  arrange(season, tmYrGameOrder) %>%
  group_by(season, team)

current <- standings %>%
  filter(tmYrGameOrder <= 8 & season == "2017/18") %>%
  select(team, tmYrGameOrder, season, cumGF) %>%
  arrange(season, tmYrGameOrder) %>%
  group_by(season, team)

manC <- standings %>%
  filter(tmYrGameOrder <= 8 &
           season == "2017/18" & team == "Man. City") %>%
  select(team, tmYrGameOrder, season, cumGF) %>%
  arrange(season, tmYrGameOrder) %>%
  group_by(season, team)

prev %>%
  plot_ly(
    x =  ~ tmYrGameOrder,
    y =  ~ cumGF,
    hoverinfo = "text",
    text =  ~ paste0(team, "<br>", season, "<br>",
                     cumGF, " goals")
  ) %>%
  add_lines(color = I("lightgrey"), name = "Prev Years") %>%
  add_lines(
    data = current,
    x =  ~ tmYrGameOrder,
    y =  ~ cumGF,
    color = I("red"),
    name = "2017/18"
  ) %>%
  add_lines(
    data = manC,
    x =  ~ tmYrGameOrder,
    y =  ~ cumGF,
    color = I("blue"),
    name = "Man City 17/18"
  ) %>%
  layout(
    title = "Goals for by team by Year",
    xaxis = list(title = "Games played"),
    yaxis = list(title = "Cumulative Goals")
  ) %>%  config(displayModeBar = F, showLink = F)


```

Everton (2005/06) now replace Palace as the lowest scorers at this stage of the season. The Toffees had finished 4th the previous campaign


***

<p class="factoid">After 8 League games, City have already scored as many goals as 16 teams have managed after 38 matches in previous EPL campaigns - including their team of 2006/07</p>

***

## Arsenal conceding Away {-}

Courtesy of a 90th minute strike by Watford's Cleverley, Arsenal(26) now only have Crystal Palace(29) ahead of them in terms of conceding the most EPL goals in 2017

```{r ARS_GA_08}

standings %>%
  filter(venue == "A") %>%
  select(gameDate, GA, team) %>%
  mutate(year = str_sub(gameDate, 1, 4)) %>%
  group_by(year, team) %>%
  summarise(tot = sum(GA)) %>%
  arrange(desc(tot)) %>%
  filter(team == "Arsenal") %>%
  plot_ly(x =  ~ year,
          y =  ~ tot,
          color = I("red")) %>%
  layout(
    title = "Arsenal goals conceded by calendar year",
    xaxis = list(title = ""),
    yaxis = list(title = "Goals conceded")
  ) %>%  config(displayModeBar = F, showLink = F)

```

1992 and 2017   (at the time of writing) are part years

Up to and including 2007, Arsenal conceded 20 away goals in only two years. Since then, they have only dipped below this level once.       
2011 included an 8-2 drubbing at Old Trafford

***

## Totally Football and Guardian Weekly Review Podcasts {-}


<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtrans`


## Final Factoid {-}

<p class="factoid">de Bruyne was withdrawn after 66 minutes with 4 assists - one shy of the EPL record</p>


---

## End of Season Update {-}

Speroni amd McAuley remained the oldest player(38.6 years) and outfielder(38.2 years) to appear this season, although both were bench players by the end of the season

#### Mo's MO {-}

The heat map gets updated and you can see that there was just one more 0-0 result, a disappointing end of calendar home game against Southampton. The most common result..

..? min table


#### City on the Rampage {-}

For the remainder of the season City were there or thereabouts as the top scorers ever after each round of games and..

```{r}


#mnc this year
mnc <- standings %>%
  filter(team == "Man. City", season == "2017/18") %>%
  select(team, season, tmYrGameOrder, cumGF, GF, GA, OppTeam)


## other teams

best <- standings %>%
  filter(tmYrGameOrder <= 39, season < "2017/18") %>%
  arrange(desc(cumGF)) %>%
  group_by(tmYrGameOrder) %>%
  slice(1) %>%
  select(team, season, cumGF, tmYrGameOrder) %>%
  ungroup() #esssential


plot_ly() %>%
  add_lines(
    data = mnc,
    x =  ~ tmYrGameOrder,
    y =  ~ cumGF,
    line = list(shape = "hv"),
    color = I("lightblue"),
    name = "Man. City",
    hoverinfo = "text",
    text = ~ paste0("Round: ", tmYrGameOrder, "<br>v ", OppTeam, "<br>",
                    GF, "-", GA)
  ) %>%
  add_lines(
    data = best,
    x =  ~ tmYrGameOrder,
    y =  ~ cumGF,
    line = list(shape = "hv"),
    color = I("red"),
    hoverinfo = "text",
    text = ~ paste0(
      "Round: ",
      tmYrGameOrder,
      "<br>",
      cumGF,
      "<br>",
      team,
      "<br>",
      season
    ),
    name = "Best"
  ) %>%
  
  layout(
    title = "Man City goalscoring over season vs best ever",
    xaxis = list(title = "Games Played"),
    yaxis = list(title = "Cumulative Points")
  ) %>% config(displayModeBar = F, showLink = F)

```


Interestingly, their 2013/14 title-winning team reached 100 goals after 37 games but only scored 2 in their final game and thus could not match the 2009/10 Chelsea side that pummeled Wigan 8-0 on the final day of that season

---

#### Arsenal conceding Away {-}

In the end, Arsenal did improve sonewhat. The 35 goals conceded was some way off Stoke's 40 and not even a club EPL record. However, the away travails continued into 2018 so that by mid-April they were the only club in the top five divisions of English football not to gain points on their travels since the turn of the year

---

## Code Hints {-}

  * plotly - heatmaps, combining two data.frames
  * base - expand.grid and t(for transform)