# Week 9 {-}



## Match of the Day{-}

Sunday games rule; as Arsenal consign Koeman to the dustbin and Klopp may start to worry


```{r topGame_09}

embed_youtube("BggJS6xh2JE", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***

## Managers Out {-}

There have already been three dismissals with less than a quarter of the season gone

1. Crystal Palace: de Boer  - not matching playing styles to player skills
2. Leicester City: Shakespeare - board were reluctant in first place
3. Everton: Koeman - poor results after big summer spending spree

Given that all the promoted clubs have performed well, several others will be nervous

1. Stoke - Mark Hughes's team are 17th and have conceded most in the division
2. West Ham - Bilic's team look disorganized. Defeat at Palace next time out could be it
3. WBA - Pulis team have 2 wins in 18 EPL games, both 1-0
4. Liverpool - Klopp's team are no better off than when Rodger's was sacked two years ago and glaring defensive         deficiencies have not been addressed

How do the dismissals compare with previous seasons
  
```{r mangagers_09}



temp <- managers %>%
  filter(is.na(Caretaker) & Left != ("2017-10-23")) %>%
  mutate(
    year = year(Left),
    month = month(Left),
    day = day(Left),
    name = paste0(FirstName, Lastname)
  ) %>%
  select(ManagerID, TeamID, name, year, month, day, Left) %>%
  filter(month %in% 8:9 | (month == 10 & day < 24)) %>%
  group_by(year) %>%
  arrange(Left) %>%
  mutate(order = row_number())

temp %>%
  count(year) %>%
  plot_ly(x =  ~ year, y =  ~ n) %>%
  add_bars(color = I("lightblue")) %>%  config(displayModeBar = F, showLink = F)

```

So higher than average but not exceptional. The most recent time when four had received their marching orders was in 2004/5. Houllier (Liverpool) and Ranieiri (Chelsea) were big names to go in the summer break and Sturrock (Southampton), Robson (Newcastle), Souness (Blackburn) and Adams (Leicester) were also out by this time of the season

The cull did not end there, as Megson (WBA), Jones (Wolves), Santini (Spurs), Redknapp( Portsmouth) and Wigley (Southampton again) were also out by Christmas. Fully half the division had new managers

***

## Player Age Spread {-}

It has been well-documented that Everton's summer spending was high on No 10's (Sigurdsson, Rooney and Klaassen) and low on goal-scorers - Lukaku and Barkley were their 1-2 scorers in each of the past two seasons

However, the age distribution of players is also out of whack

```{r age_everton_09}

ages <- summary %>%
  filter(season == "2017/18" & PLAYERID != "OWNGOAL") %>%
  mutate(age = (Sys.Date() - born) / 365.25) %>%
  select(name, TEAMNAME, mins, age)

ages %>%
  filter(TEAMNAME == "Everton") %>%
  plot_ly(
    x =  ~ age,
    y =  ~ mins,
    hoverinfo = "text",
    text =  ~ paste0(name, "<br>Age: ", round(age, 0),
                     "<br>Mins: ", mins)
  ) %>%
  layout(
    xaxis = list(title = "Age"),
    yaxis = list(title = "Minutes Played this Season")
  ) %>%  config(displayModeBar = F, showLink = F)


```
_NB This data will change as the season progresses_

They are a mix of promising youngsters and ageing veterans with only Keane - of outfield regulars (top 11 in minutes played) - in the sweet 24-27 age range

They also lead the division in regulars  who have turned 32, at the time of writing 

```{r oldies_09}

ages %>%
  arrange(desc(mins)) %>%
  group_by(TEAMNAME) %>%
  slice(1:11) %>%
  filter(age > 32) %>%
  mutate(age = round(age, 1)) %>%
  group_by(TEAMNAME) %>%
  select(player = name, team = TEAMNAME, age, mins) %>%
  
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = FALSE,
      info = FALSE
    )
  )

```


***

## Rooney of the Rovers {-}


<p class="factoid">If Rooney fails to create a goal in the first 8 minutes of his next league game, he will 
record his longest ever stretch without an assist, 1455 minutes, going back to pre-Christmas 2016</p>

On the other hand, he did score with a cracking shot from outside the area - reminiscent of his first ever-strike - also against Arsenal

```{r rooney_goal_09}

embed_youtube("E4dBWUjuCU0", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

It was his only goal against the soon-to-be-retired Seaman, but he has scored against a further 77 goalkeepers


```{r rooney_goalies_09}


## calc all goalies
goalies <- playerGame %>%
  filter(str_sub(POSITION, 1, 1) == "G" & mins > 0) %>%
  select(MATCHID, TEAMMATCHID, name, PLAYERID, gameDate, season)

# link Rooney goals to goalies he has faces, summarize and plot
playerGame %>%
  filter(mins > 0) %>%
  left_join(goalies, by = "MATCHID") %>%
  filter(TEAMMATCHID.x != TEAMMATCHID.y)  %>%
  select(
    player = name.x,
    TEAMNAME,
    Gls,
    mins,
    gameDate = gameDate.x,
    PLAYERID.x,
    PLAYERID.y,
    goalie = name.y,
    Opponents,
    season = season.x
  ) %>%
  group_by(player, PLAYERID.x, PLAYERID.y, goalie) %>%
  summarise(apps = n(),
            totMins = sum(mins),
            goals = sum(Gls)) %>%
  ungroup() %>%
  filter(PLAYERID.x == "ROONEYX") %>%
  plot_ly(
    x =  ~ totMins,
    y =  ~ goals,
    hoverinfo = "text",
    text =  ~ paste0(goalie, " (", goals, ")",
                     "<br>", totMins, " mins")
  ) %>%
  layout(
    title = "Rooney EPL goals by Opponent Goalie",
    xaxis = list(title = ("Minutes Played")),
    yaxis = list(title = "Goals scored")
  ) %>%  config(displayModeBar = F, showLink = F)


```

Hover points for details

***

## Totally Football and Guardian Weekly Review Podcasts {-}


<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6419937-koeman-and-everton-consciously-uncouple-is-bilic-next-for-the-chop/embed/v4?eid=AQAAADEy7lnh9WEA" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6420384-koeman-given-the-chop-at-everton-and-lovren-thrown-under-the-bus-football-weekly/embed/v4?eid=AQAAAAcy7lmg92EA" title="audioBoom player"></iframe>

***

## Results and Table {-}


```{r results_09}

weekly_results("2017-10-17","2017-10-24")

```


```{r standings_09}

weekly_table("2017-10-17")

```

---

## Final Factoid {-}

<p class="factoid">Manchester City already have six players with seven or more points (goals+assists). Eleven teams have none</p>


---

## End of Season Update {-}

#### Managers Out {-}

It was quite the turnover

OK, Klopp was always a long-shot and the defence improved with a steadier Karius and the purchases of Robertson and van Dilk.

However, Stoke, West Ham and WBA(2x) did soon press the eject button. Adding in Swansea and Southampton provides 9 removals in office with Wenger also ~eased out~ resigning at the end of the season


#### Player Age Spread {-}

Everton did rectify the ship via return from injuries e.g Bolasie, Coleman and winter purchases e.g Walcott and Tosun. In the end .. 10 of the 20 players who played over 500 minutes were in the 25-30 age range. However, their top-class players such as Jageilka, Rooney and Baines are all well past their best but were nevertheless regulars at season's end as the graph below indicates

```{r agesu_09}
temp <- standings %>%
  filter(season == "2017/18") %>%
  left_join(playerGame) %>%
  filter(START > 0) %>%
  group_by(team, tmYrGameOrder) %>%
  summarize(old = round(mean(age, na.rm = TRUE), 2)) %>%
  ungroup()

ars <- temp %>%
  filter(team == "Arsenal")
eve <- temp %>%
  filter(team == "Everton")
wba <- temp %>%
  filter(team == "West Brom")

temp %>%
  group_by(team) %>%
  plot_ly() %>%
  add_lines(
    x =  ~ tmYrGameOrder,
    y =  ~ old,
    color = I("lightgrey"),
    name = "Other Clubs",
    hoverinfo = "text",
    text = ~ paste0(old, "<br>",
                    team)
  ) %>%
  add_lines(
    data = ars,
    x =  ~ tmYrGameOrder,
    y =  ~ old,
    color = I("red"),
    name = "Arsenal",
    hoverinfo = "text",
    text = ~ paste0(old, "<br>",
                    team)
  ) %>%
  add_lines(
    data = eve,
    x =  ~ tmYrGameOrder,
    y =  ~ old,
    color = I("blue"),
    name = "Everton",
    hoverinfo = "text",
    text = ~ paste0(old, "<br>",
                    team)
  ) %>%
  add_lines(
    data = wba,
    x =  ~ tmYrGameOrder,
    y =  ~ old,
    color = I("black"),
    name = "WBA",
    hoverinfo = "text",
    text = ~ paste0(old, "<br>",
                    team)
  ) %>%
  layout(
    title = "Average Age of Starters by Club by Game",
    xaxis = list(title = "Game Order"),
    yaxis = list (title = "Mean Age")
  )
```

WBA were the most likely to field the oldest team but experience could not help them avoid relegation

Arsenal had the distinction of having had rounds in which they had the oldest and youngest starting lineups

---

Zabaleta was comfortably the over-30 outfielder with most minutes. His record contrast with fellow West Ham oldie, Patrice Evra whose only two starts (the second of which he  was subbed at half-time) coincided with 4-1 defeats 

---

## Rooney of the Rovers {-}

Rooney's career is petering out after a bright start to his second spell at Everton    

His assist drought ended at a whacking 1747 minutes and .. goal drought also worst of career

H has been subbed (often to his disgust) .. second in league and more than twice the rate of any other season. Still he did get at least 10 goals and 10 assists in 11 consecutive seasons so it havs been a great run

---

Manchester City attacking prowess is reflected in the five players who secured 25 or more points including deBruyne, Sterling, Aguero, David Silva and Sane. This matches the only other team which managed a fiver their own 2013/14 side - with Aguero and Siva repeat performers

---

## Code Hints {-}

 * dplyr - count()


