# Week 16 {-}


Bizarrely, the top two teams contrive to score three cock-up goals but it now looks a long way back for United to catch City


***

## City Dominance {-}

With an eleven point lead over their nearest rivals, Man City are looking home and hosed for the title. Their 46 point tally after 16 games is an EPL record, three greater than Chelsea 2005/6, but the point asvantage over second-placed, Manchester United, is actually not the greatest at this stage of a season.    
 In 1993/4, Man. Utd. held a 12 point lead over their nearest rivals, Blackburn, which had narrowed to,a still comfortable, 8 by the season end

Here are the biggest comebacks each season (hover for details)

```{r big_lead16}

leaders <- standings %>% 
  filter(position==1) %>% 
  select(leader=team,season,tmYrGameOrder,leaderPts=cumPts)

df <- standings %>% 
  filter(final_Pos==1) %>% 
  select(winner=team,season,tmYrGameOrder,winnerPts=cumPts) %>% 
  left_join(leaders,by=c("season"="season","tmYrGameOrder"="tmYrGameOrder")) %>% 
  mutate(diff=winnerPts-leaderPts) %>% 
  filter(winner!=leader) %>% 
  arrange(diff,desc(tmYrGameOrder)) %>% 
  group_by(season) %>% 
  slice(1) %>% 
  mutate(diff=if_else(diff==0,-0.2,diff)) %>% 
  mutate(diffnumber=if_else(diff==0,0,-diff)) 

legendtitle <- list(yref='paper',xref="paper",y=1.05,x=1.2,font=list(size= 14, color="#ff0000"),    text="Champions",showarrow=F)

df %>% 
    plot_ly(x=~season,y=~diff,color=~winner) %>% 
  add_bars(
           hoverinfo="text",
           text=~paste0(
             season,
             "<br>",leader," led ",winner," by ",diffnumber,' points',
             "<br>after Round ",tmYrGameOrder)
           ) %>% 
  layout(title="Maximun points deficit made up by Champion by Season",
         annotations=legendtitle,
         margin=list(b=80),
         xaxis=list(title=""),
         yaxis=list(title="Points in Arrears")) %>%  config(displayModeBar = F,showLink = F)

```

So bigger deficits later into the season have been overcome - including Manchester United over Newcastle in 1995/6 - although expectations are not high that City will implode as the Magpies did

***

City have also now recorded the highest single-season run of victories at 14

Here are the top winning runs for every team which has appeared in the Premier League. If there are ties, the first season is shown (on hover)

```{r win_runs11}

standings <- standings %>% 
  ungroup()
  

myFun <- function(x,y) {
  standings %>% 
    #ungroup() %>% 
    filter(team==x&season==y) %>% 
    arrange(tmGameOrder) %>% 
    select(res,tmGameOrder) %>% 
    mutate(cat=ifelse(res=="Win",1,0)) %>% 
    do(subSeq(.$cat)) %>% 
    filter(value==1) %>% 
    arrange(desc(slength)) %>% 
      slice(1) %>% 
      pull(slength)
}

tmYrs <-standings %>%
  select(season,team) %>%
  unique(.)


run <- map2_dbl(tmYrs$team,tmYrs$season,myFun)

df <- cbind(tmYrs,run=run)

df <-df %>% 
  group_by(team) %>% 
  arrange(desc(run), desc(team), season) %>% 
  slice(1)

df %>% 
  plot_ly(x=~run,y=~fct_reorder(team, run), color=~team, showlegend = FALSE, height = 1000) %>% 
  add_bars(hoverinfo = "text", width=0.7,
           text=~paste0(run,"<br>",season)) %>% 
  layout(title="Longest Sequences of Wins within season",
         margin=list(l=100,pad=5),
          xaxis=list(title="Wins in Row"),
         yaxis=list(title="")
) %>%  config(displayModeBar = F,showLink = F)



```

Brighton will be hoping to have got themselves off the bottom of the chart by the end of the season ( _which they did_ )


<p class="factoid">Benteke's missed penalty means he has now gone 1000 minutes without scoring in the league
</p>

***

## West Brom Woes {-}

WBA remain just above the drop zone but have not benefited from the Pardew bounce as the new management has produced no goals in two games against teams below them

They have now gone 14 games without winning.   

How often does that occur and is it a presage of relegation?


```{r winless_run16}

standings %>% 
  mutate(cat=ifelse(res=="Win",1,0)) %>% 
  select(gameDate,season,team,cat,final_Pos) %>% 
  arrange(gameDate) %>% 
  group_by(season,team) %>% 
   mutate(roll_sum = roll_sum(cat, 14, align = "right", fill = NA)) %>% 
  filter(roll_sum==0) %>% 
  ungroup() %>% 
  select(-c(gameDate,cat,roll_sum)) %>% 
  unique() %>%
 DT::datatable(width=350,class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))


```

So, this poor sequence typically occurs once a season, with WBA featuring on one previous occasion - when they escaped by the skin of their teeth. 
Generally, such a run will mean the drop but there are obviously many teams still in the mix this year and they are currently just above the trap-door


<p class="factoid">There was a greater than 20 year age gap between two outfield bench-warmers for the first time this season. WBA's Gareth McAuley(38) and Rayaan Tulloch(16) also set a club record</p>

***

## Tweet of the Week {-}

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Shawcross&#39;s feat of goal and OG in same <a href="https://twitter.com/hashtag/EPL?src=hash&amp;ref_src=twsrc%5Etfw">#EPL</a> game is 32nd time it has occurred and first by <a href="https://twitter.com/hashtag/SCFC?src=hash&amp;ref_src=twsrc%5Etfw">#SCFC</a> player. No-one has done it more than once</p>&mdash; Andrew Clark (@pssGuy) <a href="https://twitter.com/pssGuy/status/940072530953383937?ref_src=twsrc%5Etfw">December 11, 2017</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>



***

## Totally Football and Guardian Weekly Review Podcasts {-}



<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6546608-city-out-of-sight-west-ham-win-at-last-and-arsenal-off-to-ostersunds/embed/v4?eid=AQAAAHmuLlqw5GMA" title="audioBoom player"></iframe>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6546651-city-slickers-benteke-s-blunder-and-classic-allardyce-football-weekly/embed/v4?eid=AQAAAOLcLlrb5GMA" title="audioBoom player"></iframe>

***

## Results and Table {-}

```{r results_16}

weekly_results("2017-12-04","2017-12-11")

```


```{r standings_16}

weekly_table("2017-12-11")

```




## Final Factoid {-}

<p class="factoid">Palace are only the second team in EPL history to be bottom after a five-game unbeaten run (QPR 2012/13)</p>


---

## End of Season Update {-}

#### City dominance {-}

Manchester City extended their winning run to 18 before it was ended by Crystal Palace, with Liverpool inflicting the first defeat on them two games later

Brighton did manage two wins on the trot, leaving Cardiff as the only team in Premier League history never to have recorded successive victories

The most common best-sequence is five; achieved by 14 of the 49 teams to ever appear in the Premier League. Let's see how often that length run has been achieved at least in a season, by club


```{r five_winsU_16}
fiveWins <- function(x,y) {
  standings %>% 
    #ungroup() %>% 
    filter(team==x&season==y) %>% 
    arrange(tmGameOrder) %>% 
    select(res,tmGameOrder) %>% 
    mutate(cat=ifelse(res=="Win",1,0)) %>% 
    do(subSeq(.$cat)) %>% 
    filter(value==1,slength>4) %>% 
    nrow()
}


fiveYearRun <- map2_dbl(tmYrs$team,tmYrs$season,fiveWins)

## purr list col to hold years it happened
fiveYears_df <- cbind(tmYrs,count=fiveYearRun) %>% 
  filter(count>0) %>% 
  count(team, sort=TRUE)


fiveYears_df %>% 
  plot_ly(x=~n,y=~fct_reorder(team, n), color=~team, showlegend = FALSE, height = 1000) %>% 
  add_bars(hoverinfo = "text", width = 0.7,
           text=~paste0(n)) %>% 
  layout(title="Seasons with at least one 5-wins Run",
         margin=list(l=100,pad=5),
          xaxis=list(title="Number of Seasons"),
         yaxis=list(title="")
) %>%  config(displayModeBar = F,showLink = F)

```

Unsurprisingly, Manchester United top the list. The only seasons they have failed to achieve the feat are 2010/11 (when they nonetheless won the title by nine points); 2013/14 and 2015/16. It seems a bit odd to see Bolton and Leeds featuring in the top ten

---

Benteke finally ended his drought at just over 1100 minutes but he managed only three (including a mishit penalty) in over 2200 minutes (equivalent to 25 games) and ended the season on the bench . On the plus side, his nine assists was more than double that of any other of his six seasons in the league

---

#### West Brom Woes {-}

Things got little better for West Brom. The winless run extended to 20 and Pardew was dismissed before season's end. Virtually the only highlight of the season was a run of five unbeaten games at the tail-end of the season under caretaker manager, Darren Moore, including an away victory at Old Trafford 


---

## Code Hints {-}

 * plotly - annotations for legend title
 * purrr  - map2 function example
 * Rcpp   - roll_sum()
