# Week 17 {-}


Second time recently, Watford have removed Deeney late in a winning position only to end up with no points. Only 6 points separate the bottom half of the table


***

<p class="factoid">Watford were ahead of Crystal Palace for 85 minutes in Tuesday's game - 15 minutes more than Palace have led in matches the whole season to date
</p>




***

## Yet Another Man City Chart {-}

With Man. City racking up another four goals away from home, they are well on course to break Chelsea's season record of 103 in 2009/10, when Drogba (29) and Lampard (22) led the way

At the time of writing, after Round 17 of 38, they have already scored more goals than half the teams did in ANY premier league season bar 1992/3 - when 42 games were played


```{r mnc_goals_17}

mncGoals <- standings %>%
  filter(team == "Man. City",tmYrGameOrder==17) %>%
  arrange(desc(gameDate)) %>%
  slice(1) %>%
  pull(cumGF)

theTitle = glue("Teams scoring less than {mncGoals} goals in full season")

standings %>%
  filter(season < "2017/18") %>%
  group_by(team, season) %>%
  arrange(desc(gameDate)) %>%
  slice(1) %>%
  ungroup() %>%
  filter(cumGF < mncGoals) %>%
  group_by(season) %>%
  tally() %>%
  plot_ly(x =  ~ season, y =  ~ n) %>%
  add_bars(color =  ~ season, showlegend = FALSE) %>%
  layout(
    title = theTitle,
    margin = list(b = 100),
    xaxis = list(title = ""),
    yaxis = list(title = "Count")
  ) %>%  config(displayModeBar = F,showLink = F)

```

***

## Goal Scoring by Team {-}

Let's take a look at how the goalscoring has panned out by team, by minute.
You may want to zoom in at 45 and 90 minute mark as this includes any extra time - when many goals are scored

```{r GF_by_min_17}

each <- playerGame %>%
  select(PLAYER_MATCH, season, TEAMNAME, venue, Opponents) %>%
  right_join(goals) %>%
  select(season, TEAMNAME, TIME, venue, Opponents) %>%
  filter(season == "2017/18")

all <- each %>%
  group_by(TEAMNAME) %>%
  tally()


each %>%
  left_join(all) %>%
  plot_ly(x =  ~ jitter(TIME),
          y =  ~ fct_reorder(TEAMNAME, n)) %>%
  add_markers(
    hoverinfo = "text",
    color =  ~ venue,
    colors = c("blue", "orange"),
    text =  ~ paste0(TIME, " v ", Opponents)
  ) %>%
  layout(
    margin = list(l = 120),
    title = "Time of goals scored, by venue, 2017/8",
    xaxis = list(title = "Minutes Played"),
    yaxis = list(title = "")
  ) %>%  config(displayModeBar = F, showLink = F)


```


At the time of writing, points to note are that even though Manchester City have recorded 53 goals, the earliest is in the 10th minute while Liverpool's  have yet to notch one before the 15th.    

Palace have not scored away at all nor found the target between the 56th and 88th minute

***

## David Silva Acclaim {-}

Following another outstanding display against Swansea, Silva is winning a host of plaudits. However, it should be noted that many of his own team are actually putting up statistically better seasons in terms of goals and assists per 90 minutes played

```{r man_city_points_17}

playerGame %>%
  filter(
    TEAMNAME == "Man. City" &
      season == "2017/18" &
      PLAYERID != "OWNGOAL" & gameDate < as.Date("2017-12-15")
  ) %>%
  group_by(PLAYERID, name) %>%
  summarize(
    goals = sum(Gls),
    assists = sum(Assists),
    points = goals + assists,
    mins = sum(mins, na.rm = T),
    ppg = round(90 * (goals + assists) / mins, 2)
  ) %>%
  ungroup() %>%
  arrange(desc(ppg)) %>%
  filter(mins > 499) %>%
  select(` ` = name,
         goals,
         assists,
         points,
         `Pts per 90mins` = ppg) %>%
  head(10) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = FALSE,
      searching = FALSE,
      info = FALSE
    )
  )

```

 _N.B. Up to two assists are allowed per goal_

After 17 games, he is only sixth on the team for those having played at least 500 minutes

However, showing City's all-round attacking prowess, Silva's points per 90 minutes would translate to being highest-ranked on most other teams

```{r silva_points_17}

playerGame %>%
  filter(season == "2017/18" &
           PLAYERID != "OWNGOAL" & gameDate < as.Date("2017-12-15")) %>%
  group_by(PLAYERID, name, TEAMNAME) %>%
  summarize(
    goals = sum(Gls),
    assists = sum(Assists),
    points = goals + assists,
    mins = sum(mins, na.rm = T),
    ppg = round(90 * (goals + assists) / mins, 2)
  ) %>%
  ungroup() %>%
  group_by(TEAMNAME) %>%
  filter(mins > 499) %>%
  arrange(desc(ppg)) %>%
  group_by(TEAMNAME) %>%
  slice(1) %>%
  ungroup() %>%
  select(-c(PLAYERID, mins)) %>%
  arrange(desc(ppg)) %>%
  rename(` ` = name,
         `Pts per 90mins` = ppg,
         ` ` = TEAMNAME) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )
```


***


<p class="factoid">Everton have avoided defeat five times this season when they have only notched one goal. On each occasion, Rooney has been the scorer - earning the team seven points</p>

***

## Tweet of the Week {-}


<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Tough quiz<a href="https://twitter.com/hashtag/MCFC?src=hash&amp;ref_src=twsrc%5Etfw">#MCFC</a> have set the single <a href="https://twitter.com/hashtag/premierleague?src=hash&amp;ref_src=twsrc%5Etfw">#premierleague</a> season record of consecutive wins. Back in 1992/93 six clubs set their best ever single season run<br> Can you name them?<a href="https://twitter.com/BenDinnery?ref_src=twsrc%5Etfw">@BenDinnery</a> <a href="https://twitter.com/jalapic?ref_src=twsrc%5Etfw">@jalapic</a></p>&mdash; Andrew Clark (@pssGuy) <a href="https://twitter.com/pssGuy/status/940084359335309312?ref_src=twsrc%5Etfw">December 11, 2017</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


Nobody could/would!

The answer is 

* 7 Sheffield Wednesday
* 5 Crystal Palace; Norwich City
* 3 Oldham Athletic; QPR; Sheffield Utd.

***

## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6552940-recency-bias-revel-ations-and-the-rooney-contradiction/embed/v4?eid=AQAAAJunMlps_WMA" title="audioBoom player"></iframe>

***
<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6552886-manchester-city-make-history-and-arsenal-in-89-football-weekly-extra/embed/v4?eid=AQAAAG2nMlo2_WMA" title="audioBoom player"></iframe>

***

## Results and Table {-}

```{r results_17}

weekly_results("2017-12-11","2017-12-14")

```


```{r standings_17}

weekly_table("2017-12-14")

```




## Final Factoid {-}

<p class="factoid">103 players have now scored away from home - all of them totalling more than managed by the entire Palace squad</p>


---


## End of Season Update {-}

#### Yet Another Man City Chart {-}

As the season progressed, the chart above gradually filled up and became pretty meaningless

However, Man City's goal difference of 79, which is the highest in Premier League history, also exceeds the goals scored by most teams

```{r mnc_goalsu_17}

mncGD <- standings %>%
  filter(team == "Man. City") %>%
  arrange(desc(gameDate)) %>%
  slice(1) %>%
  pull(cumGD)

theTitle = glue("Teams Goals For less than Man City's {mncGD} Goal Difference ")

standings %>%
  group_by(team, season) %>%
  arrange(desc(gameDate)) %>%
  slice(1) %>%
  ungroup() %>%
  filter(cumGF < mncGD) %>%
  group_by(season) %>%
  tally() %>%
  plot_ly(x =  ~ season, y =  ~ n) %>%
  add_bars(color =  ~ season, showlegend = FALSE) %>%
  layout(
    title = theTitle,
    margin = list(b = 100),
    xaxis = list(title = ""),
    yaxis = list(title = "Count")
  ) %>%  config(displayModeBar = F, showLink = F)
```

As you can see their year ending goal difference of 79 exceeds virtually every teams goals for. This season, only Liverpool with 84 better that (and .. is best for second team?)

---

## Goal Scoring by Team {-}

Most teams score more goals in the second half of games as legs tire; teams press for a crucial goal; subs energize a team; and extra minutes are added

```{r goal_timesu_17}

df <- playerGame %>%
  select(PLAYER_MATCH, season, TEAMNAME, venue, Opponents) %>%
  right_join(goals) %>%
  select(season, TEAMNAME, TIME, venue, Opponents) %>%
  filter(season == "2017/18")


first <- df %>%
  filter(TIME <= 45) %>%
  group_by(TEAMNAME) %>%
  tally() %>%
  rename(first = n)

df %>%
  filter(TIME > 45) %>%
  group_by(TEAMNAME) %>%
  tally() %>%
  rename(second = n) %>%
  inner_join(first) %>%
  mutate(tot = first + second, pc = round(100 * second / tot, 1)) %>%
  plot_ly(x =  ~ pc , y = ~ fct_reorder(TEAMNAME, pc)) %>%
  add_bars(
    color = ~ TEAMNAME,
    hoverinfo = "text",
    text =  ~ paste0(TEAMNAME, "<br>Tot:",
                     tot, "<br>%: ", pc)
  ) %>%
  hide_legend() %>%
  layout(
    margin = list(l = 120, pad = 4),
    title = "% Goals scored in Second Half",
    xaxis = list(title = "Percentage "),
    yaxis = list(title = "")
  ) %>% config(displayModeBar = F, showLink = F)
```

Huddersfield are the only team to score the majority of their goals before half-time, and indeed did not score at all at home between the 50th and 90th minute

---

Here is the distribution over all 26 seasons

```{r GF_distributionu_17}
df_all <- playerGame %>%
  select(PLAYER_MATCH, season, TEAMNAME, venue, Opponents) %>%
  right_join(goals) %>%
  select(season, TEAMNAME, TIME, venue, Opponents)


first <- df_all %>%
  filter(TIME <= 45) %>%
  group_by(TEAMNAME, season) %>%
  tally() %>%
  rename(first = n)

all_years <- df_all %>%
  filter(TIME > 45) %>%
  group_by(TEAMNAME, season) %>%
  tally() %>%
  rename(second = n) %>%
  inner_join(first) %>%
  mutate(tot = first + second, pc = round(100 * second / tot, 1))

this_year <- all_years %>%
  filter(season == "2017/18")


all_years %>%
  plot_ly(x =  ~ pc) %>%
  add_histogram(color = I("blue"),
                opacity = 0.7,
                name = "All EPL Seasons") %>%
  add_histogram(
    data = this_year,
    x =  ~ pc,
    color = I("red"),
    name = "2017/18"
  ) %>%
  layout(
    barmode = "overlay",
    title = "Distribution of Percentage of Goals Scored Second Half",
    xaxis = list(title = "Percentage"),
    yaxis = list(title = "Count")
  )

```

The outliers are Aston Villa, 2014/15(9/31) and Reading, 2007/08(31/41)

---

#### David Silva Acclaim

If anything, his season improved but he still remained sixth on the Man City list in terms of goals and assists per game

```{r man_city_pointsu_17}

playerGame %>%
  filter(TEAMNAME == "Man. City" &
           season == "2017/18" &
           PLAYERID != "OWNGOAL") %>%   group_by(PLAYERID, name) %>%
  summarize(
    goals = sum(Gls),
    assists = sum(Assists),
    points = goals + assists,
    mins = sum(mins, na.rm = T),
    ppg = round(90 * (goals + assists) / mins, 2)
  ) %>%
  ungroup() %>%
  arrange(desc(ppg)) %>%
  filter(mins > 499) %>%
  select(` ` = name,
         goals,
         assists,
         points,
         `Pts per 90mins` = ppg) %>%
  head(10) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = FALSE,
      searching = FALSE,
      info = FALSE
    )
  )
```


--- 

Palace did, of course, improve and ended with 16 Six teams were equal or below them and they topped Harry Kane, by one

---

## Code Hints {-}

  * glue
