# Weeks 21/22 {-}


_This issue combines rounds 21 and 22 of the League_


The congested holiday season saw Man City's winning run end at 18 games but little to suggest they were anything other than champions-elect    
Meanwhile the transfer window opened on January 1st and Liverpool were soon the team making waves


***
<p class="factoid">Crystal Palace players had two players, Dann and Puncheon, suffer season-ending injuries from fouling Kevin de Bruyne</p>


## Transfers {-}

The transfer window started with a bang as Liverpool were involved in two of the highest ever transfer deals in Premier League history. Assuming Coutinho ultimately achieves the add-on targets, his fee exceeds the first 13 of the previous 15 record holding fees for outward transfers combined



```{r all_transfers_out_21}

playerClub %>%
  filter(!is.na(FEEOUT)) %>%
  arrange(LEFT) %>%
  filter(FEEOUT  > cummax(c(-Inf, head(FEEOUT,-1)))) %>%
  select(PLAYERID, TEAMID, LEFT, FEEOUT) %>%
  left_join(teamCodes) %>%
  left_join(playerGame) %>%
  select(
    player = name,
    team = TEAMNAME,
    date = LEFT,
    Fee = FEEOUT
  ) %>%
  unique() %>%
  filter(!is.na(player))  %>%
  plot_ly(x =  ~ date, y =  ~ Fee / 1000) %>%
  add_lines(line = list(shape = "hv")) %>%
  add_markers(
    x =  ~ date,
    y =  ~ Fee / 1000,
    hoverinfo = "text",
    text =  ~ paste0(player, "<br>",
                     team, "<br>",
                     date, "<br>",
                     Fee / 1000, " million<br>")
  ) %>%
  layout(
    title = "Progression in EPL Record Transfer Fees out<br>
  Zoom and Hover for details",
  showlegend = FALSE,
  xaxis = list(title = ""),
  yaxis = list(title = "Fee (Million Pounds)")
  ) %>%  config(displayModeBar = F, showLink = F)

```

Here are all the leading transactions by club. Just make your selection from the drop-down box



  
```{r by_team_21}

df <- playerClub %>%
  filter(!is.na(FEEOUT)) %>%
  arrange(LEFT)

myFun <- function(team) {
  temp <-  df %>%
    filter(TEAMID == team)
  
  temp[temp$FEEOUT > cummax(c(-Inf, head(temp$FEEOUT,-1))), ] %>%
    select(PLAYERID, TEAMID, LEFT, FEEOUT) %>%
    left_join(teamCodes) %>%
    left_join(playerGame) %>%
    select(
      player = name,
      team = TEAMNAME,
      date = LEFT,
      Fee = FEEOUT
    ) %>%
    unique() %>%
    filter(!is.na(player))
  
}

teamIDs <- teamGames %>%
  pull(TEAMID) %>%
  unique()

res <- map_df(teamIDs, myFun)

sd <- SharedData$new(res)



fs <- filter_select(
  id = "team",
  label = "team",
  sharedData = sd,
  group =  ~ team,
  allLevels = FALSE,
  multiple = FALSE
)

## this is needed as crosstalk does not work nicely with bootstrap, apparently
fs_nobootstrap <- fs

attr(fs_nobootstrap, "html_dependencies") <- Filter(function(dep) {
  dep$name != "bootstrap"
},
attr(fs_nobootstrap, "html_dependencies"))


chart <- sd %>%
  plot_ly(x =  ~ date, y =  ~ Fee / 1000) %>%
  add_lines(line = list(shape = "hv")) %>%
  add_markers(
    x =  ~ date,
    y =  ~ Fee / 1000,
    hoverinfo = "text",
    text =  ~ paste0(player, "<br>",
                     team, "<br>",
                     date, "<br>",
                     Fee / 1000, "<br>")
  ) %>%
  layout(
    title = "Progression in EPL Record Transfer Fees out by Club<br>
  Zoom and Hover for details",
  showlegend = FALSE,
  xaxis = list(title = ""),
  yaxis = list(title = "Fee (Million Pounds)")
  ) %>%  config(displayModeBar = F, showLink = F)   %>%
  config(displayModeBar = F, showLink = F)


tagList(fs_nobootstrap,
        br(),
        chart)

```
  
 
 

***
<p class="factoid">A week after highlighting how Man City had failed to score an early goal, they net within 40 seconds. Sterling now has the earliest and latest goals scored in league this season</p>
***




## Goals conceded {-}

A few weeks ago I created a plot that showed the time in a game at which teams had scored all of their goals. Of course, it is easy to amend the code to show when teams have conceded goals. Manchester City lead the league in both categories this season but are unlikely to match over the whole campaign the meagre 15 goals that Chelsea conceded in winning the title in 2004/5

For a change, I have added a range-selector - but you can still zoom in, if preferred

```{r GA_21}

each <- playerGame %>%
  select(PLAYER_MATCH, season, TEAMNAME, venue, Opponents) %>%
  right_join(goals) %>%
  select(season, TEAMNAME, TIME, venue, Opponents) %>%
  mutate(venue = ifelse(venue == "A", "H", "A")) %>%  ## need to reverse venues
  filter(season == "2004/05")

all <- each %>%
  group_by(Opponents) %>%
  tally()


each %>%
  left_join(all) %>%
  plot_ly(x =  ~ jitter(TIME),
          y =  ~ fct_reorder(Opponents, n)) %>%
  add_markers(
    hoverinfo = "text",
    color =  ~ venue,
    colors = c("orange", "blue"),
    text =  ~ paste0(TIME, " v ", TEAMNAME)
  ) %>%
  layout(
    margin = list(l = 120),
    title = "Time of goals conceded, by venue, 2004/05",
    xaxis = list(title = "Minutes Played", rangeslider = list(type = "date")),
    yaxis = list(title = "")
  ) %>%  config(displayModeBar = F, showLink = F)

```

A couple of trivia for Chelsea's title-winning season
  
   * Only four goals were scored in the first 2 minutes of any game, and Chelsea allowed two of them
   * They only gave up one goal in the final 20 minutes (plus added time) allowing Bolton to come back from a 2 goal deficit away from home to tie the game



***

## Tweet of the Week {-}

It was a tweet-free vacation but you can always catch my [feed here](https://twitter.com/pssGuy)

```{r}  
#shortcode("tweet", "") 
```



***

## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6579674-the-best-game-of-the-year-fa-cup-weekend-and-mark-hughes-hangs-on/embed/v4?eid=AQAAAGweUFraZWQA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6580011-wilshere-s-left-foot-why-liverpool-should-sell-coutinho-and-2018-so-far-football-weekly-extra/embed/v4?eid=AQAAAA8eUForZ2QA" title="Audioboom player"></iframe>

***

## Results and Table {-}

```{r results_21}

weekly_results("2017-12-29","2018-01-05")

```

---


```{r standings_21}

weekly_table("2018-01-05")

```




## Final Factoid {-}

<p class="factoid">Manchester City are the only team in EPL history achieve a 50+ goal difference after 22 games.
They also hold the other podium positions with 42 (2011/12) and 38(2013/14)</p>


---

## End of Season Update {-}

#### Transfers {-}

Let's see how the January transfers in have actually performed

```{r transfersu_21}

signings <- playerClub %>% 
  filter(JOINED>"2017-12-31") %>% 
  pull(PLAYER_TEAM) #47 I have 49 in need checking later


sort(names(playerGame))

chuck <- c("Joe Rodon","James Daly","Trevor Chalobah","Curtis Jones")

playerGame %>% 
  filter(PLAYER_TEAM %in% signings) %>% 
  group_by(TEAMNAME,name,PLAYERID,joined,FEE,PERMANENT) %>% 
  summarize(mins=sum(mins),goals=sum(Gls),assists=sum(Assists),age=round(max(age,na.rm=T),0)) %>%
  ungroup() %>% 
  mutate(fee=ifelse(FEE==99,"Unknown",as.character(FEE))) %>% 
  mutate(loan=ifelse(PERMANENT==0,"Yes","")) %>% 
  select(team=TEAMNAME,player=name,age,fee,loan,mins,goals,assists) %>% 
  filter(!player %in% chuck) %>% 
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))


```

A bit of a mixed bag. Arsenal's are expensive, have performed but did not contribute an end-product for the club this year. Palace's underwhelmed (3 of them never played a minute) but were mainly cover or bought for the future and they avoided relegation anyway. 

Newcastle's loans were very succesful but WBA's were not. Moura gave the impression of another expensive squad player at Spurs until an out-of-the-blue 4 assists in the final game of the season

---

#### Goals Conceded {-}

Here are this season's figures, by minute

```{r}
each <- playerGame %>% 
  select(PLAYER_MATCH,season,TEAMNAME,venue,Opponents) %>% 
  right_join(goals) %>% 
  select(season,TEAMNAME,TIME,venue,Opponents) %>% 
  mutate(venue=ifelse(venue=="A","H","A")) %>%  ## need to reverse venues 
  filter(season=="2017/18")

all <- each %>% 
  group_by(Opponents) %>% 
  tally()


each %>% 
  left_join(all) %>% 
  plot_ly(x=~jitter(TIME),y=~fct_reorder(Opponents, n)) %>% 
  add_markers(hoverinfo="text", color=~venue, colors=c("orange","blue"),
              text=~paste0(TIME," v ",TEAMNAME)) %>% 
  layout(margin=list(l=120, pad=3),autosize = F, height = 500,
         title="Time of goals conceded, by venue, 2017/18",
         xaxis=list(title="Minutes Played"),
         yaxis=list(title="")
  ) %>%  config(displayModeBar = F,showLink = F)

```


The earliest Manchester City allowed a goal at home was the 24th minute and it mattered not a jot. Leicester's goal was only an equalizer and they were then hit by a 4 goal onslaught from Aguero, one of only two such hauls all season (Salah v Watford)
 Leicester were susceptible to allowing early goals on their travels with 7 conceded in the first nine minutes of games


## Code Hints {-}

  * plotly - rangeslider
