# Week 26 {-}

## Match of the Day {-}


This week's crisis team, Chelsea.  

```{r topGame_26}

embed_youtube("MT6riYHiNUQ", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***
<p class="factoid">Jordon Ibe has picked up 3 assists in his most recent two games. He managed just one in 25 appearances for Bournemouth last season</p>

***

## Win Percentages {-}

in spite of spending well in excess of 200 million pounds on players since last season's title-winning season, Chelsea are in some danger of failing to make the top 4 and the end of Conte's reign as manager now appears to be a matter of when rather than if

He and Guardiola have now managed 64 games in the Premier League with each winning 45

Here are how the top manager/team combinations all-time(min 6 games) have started their incumbencies

```{r win_pc_26}

top5 <- managerGame %>%
  mutate(cat = ifelse(res == "Win", 1, 0)) %>%
  select(name, TEAMNAME, res, cat) %>%
  group_by(name, TEAMNAME) %>%
  summarize(
    games = n(),
    wins = sum(cat),
    win_pc = round(100 * wins / games, 2)
  ) %>%
  filter(games > 63) %>%
  arrange(desc(win_pc)) %>%
  head(5)

test <- top5 %>%
  left_join(managerGame) %>%
  select(name, TEAMNAME, res, gameDate) %>%
  mutate(cat = ifelse(res == "Win", 1, 0)) %>%
  group_by(name, TEAMNAME) %>%
  arrange(gameDate) %>%
  mutate(
    gameOrder = row_number(),
    cumwins = cumsum(cat),
    win_pc = round(100 * cumwins / gameOrder, 2)
  )


test %>%
  filter(gameOrder < 65) %>%
  mutate(manTeam = glue("{name} ({TEAMNAME})")) %>%
  plot_ly(x =  ~ gameOrder,
          y =  ~ win_pc,
          color =  ~ manTeam) %>%
  add_lines() %>%
  layout(
    title = "Leading Managers Cumulative Win %",
    xaxis = list(title = "Games Managed"),
    yaxis = list(title = "% Wins")
  ) %>%  config(displayModeBar = F, showLink = F)



```

At the 20 game mark, Conte had an 80% win rate but this has been on the slide in contrast to Guardiola's upward trend - particularly marked at the beginning of this season.
  However, Mourinho's 78% trumps both over the first seaon and two-thirds in the Premier League 


***

<p class="factoid">Benteke may be struggling to score, but at least he has 7 assists - with 4 the most he has recorded in any previous season</p>
***

## Consecutive thumpings {-}

A late collapse to Watford consigned Chelsea to consecutive losses by three goal margins for the first time since they were beaten 1-4 by Man Utd. and 3-0 by Blackburn, two powerhouses in October 1995

Here are the most recent occurrences for the 38 teams that have suffered the indignation at least once

```{r big_loss_26}

standings %>%
  select(team, gameDate, OppTeam, GF, GA) %>%
  group_by(team) %>%
  arrange(desc(gameDate)) %>%
  mutate(
    cat = ifelse(GA - GF > 2, 1, 0),
    leadCat = lead(cat),
    consec = cat + leadCat
  ) %>%
  filter(consec == 2) %>%
  group_by(team) %>%
  slice(1) %>%
  arrange(desc(gameDate)) %>%
  select(date = gameDate, team, GF, GA, opponents = OppTeam) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

```
<p>

Six teams have already had consecutive drubbings handed to them this season but Premier League ever-presents, Arsenal and Liverpool, have never suffered the fate


***

## Tweet of the Week {-}



```{r}  
shortcode("tweet", "962020532836167680") 
```


## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6641237-karma-mystery-tours-and-the-lepers-v-the-scoundrels/embed/v4?eid=AQAAAOCRfVpVVmUA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6641577-magical-mo-salah-marvellous-misses-and-celebrity-barracking-football-weekly/embed/v4?eid=AQAAAJuRfVqpV2UA" title="Audioboom player"></iframe>

***

## Results and Table {-}

```{r results_26}

weekly_results("2018-02-02","2018-02-06")

```

<p>

```{r standings_26}

weekly_table("2018-02-06")

```


## Final Factoid {-}

<p class="factoid">Looking like a 3-way race for Player of Season with pre-season favourite de Bruyne(7 goals, 22 assists) being challenged by Salah(21/11) and Kane(22/8)</p>


---


## End of Season Update {-}

Ibe had a much better season if not fully justifying his 15 millipn pound transfer fee. He lead the team in assists, with 8 ..


####  Win Percentages {-}

Extending to the full season

```{r win_pcu_26}

top5 <- managerGame %>%
  mutate(cat = ifelse(res == "Win", 1, 0)) %>%
  select(name, TEAMNAME, res, cat) %>%
  group_by(name, TEAMNAME) %>%
  summarize(
    games = n(),
    wins = sum(cat),
    win_pc = round(100 * wins / games, 2)
  ) %>%
  filter(games > 63) %>%
  arrange(desc(win_pc)) %>%
  head(5)

test <- top5 %>%
  left_join(managerGame) %>%
  select(name, TEAMNAME, res, gameDate) %>%
  mutate(cat = ifelse(res == "Win", 1, 0)) %>%
  group_by(name, TEAMNAME) %>%
  arrange(gameDate) %>%
  mutate(
    gameOrder = row_number(),
    cumwins = cumsum(cat),
    win_pc = round(100 * cumwins / gameOrder, 2)
  )


test %>%
  filter(gameOrder < 77) %>%
  mutate(manTeam = glue("{name} ({TEAMNAME})")) %>%
  plot_ly(x =  ~ gameOrder,
          y =  ~ win_pc,
          color =  ~ manTeam) %>%
  add_lines() %>%
  layout(
    title = "Leading Managers Cumulative Win %",
    xaxis = list(title = "Games Managed"),
    yaxis = list(title = "% Wins")
  ) %>%  config(displayModeBar = F, showLink = F)
```

Guardiola overtook Conte aas the season wounded down but both trail Mourinho's initial two years in the league when he won back-to-back titles. The others had to settle for one each

---

#### Consecutive thumpings {-}

As can be seen from the table above, Chelsea were not the last side to suffer the indinity this season. West Ham even made it three heavy defeats in a row

This is a fairly rare occurrence, which actually only happened once in the first 15 years of the Premier League


```{r big_lossu_26}

standings %>%
  mutate(cat = ifelse((GA - GF) > 2, 1, 0)) %>%
  arrange(gameDate) %>%
  group_by(team) %>%
  mutate(roll_sum = roll_sum(cat, 3, align = "right", fill = NA)) %>%
  filter(roll_sum == 3) %>%
  ungroup() %>%
  mutate(final = glue("{GF}-{GA} v {OppTeam}")) %>%
  select(team, season, `final game` = final, `finished` = final_Pos) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = FALSE,
      searching = FALSE,
      info = FALSE
    )
  )

```

What is even odder, is that it happened to both Newcastle and Stoke twice in the same season; yet they each finished the year in the top half of the table

At the time Newcastle started the first run with a shattering 3-0 home loss in the derby against 17th placed Sunderland, they were in 8th spot, 10 points clear of 10th. The remainder of the season they showed relegation form and in spite of finishing 9th had a -16 goal difference

Two years later after 22 games, Stoke were in 7th spot just six points off the top four. Their collapse was not so calamitous (other than the big six defeats they picked up 18 points in ten games) but like Newcastle they finshed ninth with a large GD, -14

---

Salah finished the season very strongly and desevedly took home the individual honours

---

## Code Hints {-}


 * none this week


