# Week 27 {-}

Only six away goals this week with Salah, predictably, amongst the scorers. Fourth game in a row in which they have either won or lost by three clear goals

***

## Aguero-a-go-go {-}

With the signing of Jesus last season and the pursuit of Alexis Sanchez in the recent transfer window, it appears as though Sergio Aguero is no longer the favourite son at Manchester City. However, he continues scoring at as fast a rate as ever
 His four goals against Leicester made him the first player to manage that feat on three occasions in the EPL and he now accounts for two of the four occasions in which as many goals have been scored by one player in the second forty-five minutes
 
 At the time of writing, multi-goal games now account for over 50%  of his goals - only exceeded by Kane and Henry of those players that have scored at least 100
 Hover any point for furter information
 
 
```{r multi_goals_27}

allGoals <- playerGame %>%
  group_by(PLAYERID) %>%
  filter(PLAYERID != "OWNGOAL") %>%
  summarize(totGoals = sum(Gls)) %>%
  arrange(desc(totGoals))

multiGoals <- playerGame %>%
  group_by(PLAYERID) %>%
  filter(PLAYERID != "OWNGOAL" & Gls > 1) %>%
  summarize(totMultiGoals = sum(Gls)) %>%
  arrange(desc(totMultiGoals))

test <- allGoals %>%
  left_join(multiGoals) %>%
  mutate(multi_pc = round(100 * totMultiGoals / totGoals, 1)) %>%
  filter(totGoals != 0) %>%
  mutate(multi_pc = ifelse(is.na(multi_pc), 0, multi_pc))

pgMini <-
  playerGame %>%  ## so wil only show those that have made an appearance - but that is prob ok
  select(PLAYERID, name) %>%
  unique()

df <- test %>%
  left_join(pgMini)

aguero <- df %>%
  filter(PLAYERID == "AGUEROS")

kane <- df %>%
  filter(PLAYERID == "KANEH")



plot_ly() %>%
  add_markers(
    data = df,
    x =  ~ totGoals,
    y =  ~ multi_pc,
    hoverinfo = "text",
    color = I("lightgrey"),
    name = "Other players",
    text =  ~ paste0(name, "<br>Tot: ", totGoals,
                     "<br>Multi-Game: ", multi_pc, "%")
  ) %>%
  add_markers(
    data = aguero,
    x =  ~ totGoals,
    y =  ~ multi_pc,
    color = I("blue"),
    name = "Aguero",
    hoverinfo = "text",
    text =  ~ paste0(name, "<br>Tot: ", totGoals,
                     "<br>Multi-Game: ", multi_pc, "%")
  ) %>%
  
  add_markers(
    data = kane,
    x =  ~ totGoals,
    y =  ~ multi_pc,
    color = I("red"),
    name = "Kane",
    hoverinfo = "text",
    text =  ~ paste0(name, "<br>Tot: ", totGoals,
                     "<br>Multi-Game: ", multi_pc, "%")
  ) %>%
  layout(
    title = "Share of Goals in Multi-Goal games",
    xaxis = list(title = "Goals Scored"),
    yaxis = list(title = "%")
  ) %>%  config(displayModeBar = F, showLink = F)

```
 
 It will be interesting to see if Kane keeps this up - it would help if he did not miss penalties!

## Player of Year {-}

With more than 70% of the season now under the belt, thoughts turn toward the individual prize of player of the season. Attacking players usually take the laurels (last season's awards to Kante proving an exceptiion) and that is likely to be the case again this season

Of pre-season favourites, Pogba and Sanchez  can surely be discounted whilst Sterling and Salah have forced their way into the reckoning. The general pick was de Bruyne and he has done little to disappoint

```{r}

top5 <- playerGame %>%
  filter(season == "2017/18" & gameDate < "2018-02-13") %>%
  group_by(name, PLAYERID) %>%
  mutate(points = Gls + Assists) %>%
  summarize(cat = sum(points)) %>%
  arrange(desc(cat)) %>%
  head(5) %>%
  pull(PLAYERID)

playerGame %>%
  filter(season == "2017/18" & PLAYERID %in% top5) %>%
  arrange(gameDate) %>%
  group_by(name, PLAYERID) %>%
  mutate(
    cumGoals = cumsum(Gls),
    cumAssists = cumsum(Assists),
    cumPoints = cumGoals + cumAssists
  ) %>%
  select(name,
         Goals = cumGoals,
         Assists = cumAssists,
         Points = cumPoints,
         gameDate) %>%
  plot_ly(x =  ~ gameDate,
          y =  ~ Points,
          color =  ~ name) %>%
  add_lines(
    line = list(shape = "hv"),
    hoverinfo = "text",
    text =  ~ paste0(name,
                     "<br>Goals: ", Goals,
                     "<br>assists: ", Assists)
  ) %>%
  layout(
    title = "Goals and Assist by top 5 players",
    xaxis = list(title = ""),
    yaxis = list(title = "Cumulative Points")
  ) %>%  config(displayModeBar = F, showLink = F)
```

_N.B. Up to 2 assists per goal are allowed_



Clicking on an individual player will help identify their performance over time

Aguero had a fast start with 6 goals and 4 assists in his first half-dozen games but 
was then rotated in and out of the team. However a trio of three-point games since mid-December have put him back in the race

Obviously, this could go down to the wire, but with City almost certain to win the title, the influential de Bruyne probably still has the edge


***

<p class="factoid">Nine teams won their matches in this round, but none of them have a sequence any longer than one</p>
***

## Tweet of the Week {-} 

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Chelsea have just suffered their first consecutive three goal margin defeats in <a href="https://twitter.com/hashtag/premierleague?src=hash&amp;ref_src=twsrc%5Etfw">#premierleague</a> since 1995. But which <a href="https://twitter.com/hashtag/EPL?src=hash&amp;ref_src=twsrc%5Etfw">#EPL</a> ever-presents have never had such back to back drubbings?<a href="https://twitter.com/hashtag/AFC?src=hash&amp;ref_src=twsrc%5Etfw">#AFC</a> <a href="https://twitter.com/hashtag/CFC?src=hash&amp;ref_src=twsrc%5Etfw">#CFC</a> <a href="https://twitter.com/hashtag/LFC?src=hash&amp;ref_src=twsrc%5Etfw">#LFC</a> <a href="https://twitter.com/hashtag/MUFC?src=hash&amp;ref_src=twsrc%5Etfw">#MUFC</a> <a href="https://twitter.com/hashtag/THFC?src=hash&amp;ref_src=twsrc%5Etfw">#THFC</a><a href="https://twitter.com/BenDinnery?ref_src=twsrc%5Etfw">@BenDinnery</a></p>&mdash; Andrew Clark (@pssGuy) <a href="https://twitter.com/pssGuy/status/962257142525939712?ref_src=twsrc%5Etfw">February 10, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6670169-tottenham-take-charge-super-salah-and-derry-goes-football-weekly/embed/v4?eid=AQAAAKKzglpZx2UA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6669885-the-mousa-dembele-appreciation-society-and-jese-s-hot-breath-on-charlie-adam-s-neck/embed/v4?eid=AQAAAKG1glo9xmUA" title="Audioboom player"></iframe>

***

## Results and Table {-}

```{r results_27}

weekly_results("2018-02-07","2018-02-14")

```

<p>
---

```{r standings_27}

weekly_table("2018-02-14")

```



## Final Factoid {-}

<p class="factoid">Dembele gets a lot of love in the Totally Football podcast above, but his assist against Arsenal was just the second of season and he has no goals. At 30, Spurs really need to upgrade on him and Sissoko if they want to press on next year </p>


---

## End of Season Update {-}

## Aguero-a-go-go {-}

Aguero finished the season just above the 50% mark and reached the 20 goal mark for the fourth year in succession, finally making the PFA team of the season

Kane's multi-goal scoring explots dropped a bit although he finished the season with a brace

```{r}


goals %>%
  left_join(playerGame) %>%
  filter(PLAYERID == "KANEH") %>%
  group_by(gameDate) %>%
  arrange(TIME) %>%
  mutate(order = as.factor(row_number())) %>%
  select(gameDate, TIME, order, plGameOrderApp, Opponents) %>%
  plot_ly(y =  ~ plGameOrderApp, x =  ~ TIME) %>%
  add_markers(
    color = ~ order,
    colors = c("orange", "blue", "red", "black"),
    hoverinfo = "text",
    text =  ~ paste0(
      "Game: ",
      plGameOrderApp,
      "<br>",
      Opponents,
      "<br>",
      gameDate,
      "<br>",
      TIME,
      " mins"
    )
  )  %>%
  layout(
    title = "Times at which Harry Kane scored",
    yaxis = list(title = "Game Order"),
    xaxis = list(title = "Minutes Played")
  ) %>%
  config(displayModeBar = F, showLink = F)

```

---









