# Week 29 {-}


Great come-back by Man United, who are now nine points clear of fifth-placed Chelsea and looking good for CL football again next year
Arsenal lose three league games in eight days


***
<p class="factoid">In 273 league appearances with Manchester United Patrice Evra was part of a line up that conceded 4 or more goals on 5 occasions (1.8%). For West Ham, he is 2 for 2 (100%)</p>

***

## Rod off Burnley's back {-}

As many of you have heard, Burnley finally won a Premier League match under Sean Dyche after going behind. Unsurprisingly, it was Everton - with their poor away record - that were the collaborators, allowing Chris Wood to be unguarded in the six yard area from a corner

How does this run compare with the worst ever by team?
 
 
```{r behind_29}


everBehind <- scoreLines %>%
  filter(diff == -1) %>%
  # group_by(TEAMMATCHID) %>%
  select(TEAMMATCHID, MATCHID) %>%
  unique() %>%   #10499
  left_join(standings) %>%
  select(team, res, gameDate)

everBehind %>%
  mutate(cat = ifelse(res == "Win", 1, 0)) %>%
  arrange(desc(gameDate)) %>%
  group_by(team) %>%
  do(subSeq(.$cat)) %>%
  filter(value == 0) %>%
  arrange(desc(slength)) %>%
  group_by(team) %>%
  slice(1) %>%
  arrange(desc(slength), team) %>%
  select(team, max_run = slength) %>%
  DT::datatable(
    width = 200,
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

  
```
 
 So Burnley's record does crack the top ten but is by no means the worst. Little surprise to see Sunderland and Derby County leading the field - even if over 100 games, for each, seems extraordinary
 The Man. City run ended in 2003, a period before the big investment
 

## Quick off the mark {-}

With the first goal so critical, it is interesting to see who turns the ball in first in most games. Here are all the candidates, this season

```{r quick_29}

firstGoals <- goals %>%
  left_join(matchTeam, by = c("TEAMMATCHID")) %>%
  arrange(TIME) %>%
  group_by(MATCHID) %>%
  slice(1) %>%
  select(PLAYER_MATCH, TIME, TEAMMATCHID) %>%
  left_join(playerGame) %>%
  ungroup() %>%
  filter(season == "2017/18") %>%
  select(name, TIME, TEAMNAME, Opponents, gameDate) %>%
  count(name, TEAMNAME, sort = TRUE) %>%
  rename(count = n,
         team = TEAMNAME,
         player = name)

firstGoals %>%
  arrange(desc(count), team) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )


```

Not too many surprises at the top of the list. For a full 38 game season, the record holder is a surprse: Darren Bent for Sunderland in 2009/10 when 14 of his 24 goals were the first of the game

---

Here are the leaders, by club, this year

```{r quick_leaders_29}

firstGoals %>%
  group_by(team) %>%
  arrange(desc(count)) %>%
  slice(1) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

```

_N.B. Ties for top position are not shown_

At the time of writing, Benteke tops the Crystal Palace ranking with the only two goals he has scored all season ( _Subsequently Milivojevic and Zaha had three openers_ )



***



## Tweet of the Week {-}

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Bournemouth become latest <a href="https://twitter.com/hashtag/premierleague?src=hash&amp;ref_src=twsrc%5Etfw">#premierleague</a> team to notch a run of 10 consecutive games with a goal. Unlikely that long time failures <a href="https://twitter.com/hashtag/SCFC?src=hash&amp;ref_src=twsrc%5Etfw">#SCFC</a> and <a href="https://twitter.com/hashtag/CPFC?src=hash&amp;ref_src=twsrc%5Etfw">#CPFC</a> will join them this year <a href="https://twitter.com/hashtag/AFCB?src=hash&amp;ref_src=twsrc%5Etfw">#AFCB</a> <a href="https://t.co/DNX7cOwXD5">pic.twitter.com/DNX7cOwXD5</a></p>&mdash; Andrew Clark (@pssGuy) <a href="https://twitter.com/pssGuy/status/970276394851880961?ref_src=twsrc%5Etfw">March 4, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6706745-rubbernecking-chocolate-legs-and-jay-bothroyd/embed/v4?eid=AQAAAEiMnlo5VmYA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6706857-chelsea-walk-into-city-s-jaws-arsenal-fall-again-and-the-pulis-connection-football-weekly/embed/v4?eid=AQAAAG6MnlqpVmYA" title="Audioboom player"></iframe>

***

## Results and Table {-}


```{r results_29}

weekly_results("2018-03-01","2018-03-06")

```

<p>
---

```{r standings_29}

weekly_table("2018-03-06")

```


## Final Factoid {-}

<p class="factoid">Dembele gets a lot of love in the Totally Football podcast above, but his assist against Arsenal was just the second of season and he has no goals. At 30, Spurs really need to upgrade on him and Sissoko if they want to press on next year </p>


---


## End of Season Update {-}

Amazingly, Evra had one more start for West Ham. It too ended in a 4-1 defeat to Manchester City

---


#### Rod off Burnley's back {-}

Here's how all the team's fared after going behind at least once in a game

```{r behindu_29}

scoreLines %>%
  filter(diff == -1) %>%
  # group_by(TEAMMATCHID) %>%
  select(TEAMMATCHID, MATCHID) %>%
  unique() %>%   #10499
  left_join(standings) %>%
  filter(season == "2017/18") %>%
  ungroup() %>%
  select(team, res) %>%
  group_by(team) %>%
  count(res)  %>%
  spread(res, n, fill = 0) %>%
  mutate(G = Win + Draw + Loss, ppg = round((3 * Win + Draw) / G, 2)) %>%
  select(team, G, Win, Draw, Loss, ppg) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )
```

If Huddersfield, WBA or Brighton went into arrears they could pretty well say goodbye to any points

 Arsenal actually went into arrears in games more often than Brighton, though their superior talent did ensure they more often made amends

---

####  Quick off the mark {-}

Own Goals have been beneficial to Huddersfield but two 'first-goals' have also been scored by Mooy, van la Parra and Ince- one of whom's was a last minute winnner against Watford

---

Dembele upped his game a bit at the end of the year and ended with 4 assists, good for 8th= on the team. No goals, though, and rumoured to be available this summer

---


## Code Hints {-}

 * tidyr - spread()