# Week 30 {-}


## Match of the Day {-}

Rashford schools Alex-Arnold


```{r topGame_30}

embed_youtube("CvZuoNWASbw", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***
<p class="factoid">Palace joined WBA with a league-leading ninth one-goal defeat. Every team has suffered at least one such occurrence this season. Surprisingly Huddersfield have lost by a single goal just once in their 15 defeats</p>

***

## Manager changes {-}

Several teams have changed manager's this season in an attempt to acquire a 'New manger Bounce'. Whether, or not,this concept has any validity (Managers are usually sacked after a bad run and some reversion to mean performance would be expected if they stayed) there has certainly been some significant variation in outcome

Other than Southampton - who may stll decide to take the plunge - and the promoted clubs, all those teams currently in most imminent danger have called in a new guy. 

Here are their records
 
 
```{r managers_30}

# teams that have changed managers during season
changeTeams <- managers %>%
  filter(Joined > "2017-08-10") %>%
  pull(TEAMNAME) %>%
  unique()

allManagerStints <-
  managerGame %>%
  select(name, ManagerTeam, Joined, Left) %>%
  unique()


temp <- managerGame %>%
  filter(gameDate > "2017-08-01" & TEAMNAME %in% changeTeams) %>%
  group_by(TEAMNAME, ManagerID, ManagerTeam, name) %>%
  dplyr::summarise(
    sumPoints = sum(points),
    games = n(),
    ppg = round(sumPoints / games, 2)
  ) %>%
  ungroup()
## unsworth is in their twice error in raw material
## temp just gives summary not game duration


teamRecord <- temp %>%
  inner_join(allManagerStints) %>%
  mutate(started = ifelse(
    Joined < "2017-08-10",
    as.Date("2017-08-10", origin = "1900-01-01"),
    as.Date(Joined, origin = "1900-01-01")
  )) %>%
  mutate(st = as.Date(started, origin = "1970-01-01"))
# hack but looks good hrbrmisc may proves usefurl - or lubridate

minY <- min(teamRecord$ppg) - 0.1
maxY <- max(teamRecord$ppg) + 0.1

plot_ly(teamRecord, color = ~ TEAMNAME) %>%
  add_segments(
    x = ~ st,
    xend = ~ Left,
    y = ~ ppg,
    yend = ~ ppg,
    hoverinfo = "text",
    text =  ~ paste0(name,
                     "<br>", TEAMNAME,
                     "<br>Games:", games,
                     "<br>ppg:", ppg)
  ) %>%
  layout(
    title = "Average Points per Game by Manager 2017/18",
    xaxis = list(title = ""),
    yaxis = list(title = "Points per Game")
  ) %>%  config(displayModeBar = F, showLink = F)

```
 
 Double-Click on the team name to isolate just their data. The short segments (barring de Boer at Crustal Palace) represent caretakers    
 
 Other than WBA - where Alan Pardew's managerial career in the Premier League looks set to end - there have been some improvement; Though Stoke's is marginal and may not suffice to save them
 

## Super Subs {-}

Burnley have proved this season's  good-news story and after a sticky spell are again pushing for a Europa spot. One reason has been the form of new-purchase, Chris Wood, and he came on as a sub, Sturday, to set up the first goal and score two others. Three points in a game for a sub is a pretty rare event - occurring on average just over once  a  season



```{r super_subs_30}

playerGame %>%
  filter(subOn > 0 & (Gls + Assists) > 2) %>%
  select(name, TEAMNAME, season, gameDate, Assists, Gls, mins) %>%
  arrange(desc(gameDate)) %>%
  select(
    player = name,
    team = TEAMNAME,
    season,
    date = gameDate,
    Goals = Gls,
    Assists,
    mins = mins
  ) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )


```

_N.B. Unlike official statistics, I allow up to two assists per goal_    

Oscar and David Beckham, feature twice

Stand-out performances include Solskjaer's four goals in 22 minutes and 
Wigan;s Amr Zaki who pitched in 2 goals and an assist in just 6 minutes in a game versus Hull - a team he was subsequently played for without scoring or assisting on a single goal
---



***

## Tweet of the Week {-}


```{r}  
shortcode("tweet", "973146759135481856") 
```


## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6718931-toxicity-complicity-and-pitchside-relief/embed/v4?eid=AQAAAM4NqFrThWYA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6719096-did-upton-park-die-for-this-football-weekly/embed/v4?eid=AQAAAJwNqFp4hmYA" title="Audioboom player"></iframe>

***

## Results and Table {-}


```{r results_30}

weekly_results("2018-03-06","2018-03-13")

```

<p>
---

```{r standings_30}

weekly_table("2018-03-13")

```




## Final Factoid {-}

<p class="factoid">West Ham suffer three consecutive defeats by 3 goals exactly ten years since they lost 3 games on the trot 0-4</p>


---


## End of Season Update {-}


In the end, Huddersfield did lose another game by the odd goal, 0-1 to Newcastle, but the distribution still looks odd

```{r}
standings %>%
  filter(season == "2017/18" & team == "Huddersfield") %>%
  mutate(GD = GF - GA) %>%
  plot_ly(x =  ~ GD) %>%
  add_histogram(color = "yellow")
```

Excluding Huddersfield, defeated teams lost by 1 goal .. % of the time. WBA ended up with most single game defeats

---

#### Manager changes {-}

In most seasons 38 points, a point a game, is enough to stay up, but teams worried about relegation and swapping managers are often in a worse situation  and will expect more
  From the chart above, ... might add something when cleare picture
  
  
#### Super Subs {-}

Here is the attacking impact of subs this season, by club
The 'ppg' column is an estimated per 90 minutes (but may be overstated as extra time is not included in the minutes accrued figure)


```{r super_subsu_30}

playerGame %>%
  filter(subOn > 0, season == "2017/18") %>%
  group_by(TEAMNAME) %>%
  summarize(
    totG = sum(Gls),
    totA = sum(Assists),
    totP = totG + totA,
    totM = sum(mins),
    ppg = round(100 * totP / totM, 2)
  ) %>%
  arrange(desc(totP)) %>%
  select(
    team = TEAMNAME,
    goals = totG,
    assists = totA,
    points = totP,
    mins = totM,
    ppg
  ) %>%
  
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

  
  
```

At Watford, Pereyra and Deeney have been particularly impactful, with Wood and Vokes contributing significantly at Burnley.

Tottenham trail the other top clubs, with only single strikes from Wanyama and Sissoko providing goals from off the bench

---

The Burnley defeat may have been the end of West Ham's big losses but by the end of the year they had suffered 10 three goal or larger losses, matching the 38 game season record of Derby County in 2007-8


## Code Hints {-}

  * plotly - add_segments()







