# Week 31 {-}



Abbreviated version with only four games at the week-end but Salah runs riot


***
<p class="factoid">Tenth time a Liverpool player has added an assist to a hat-trick and Salah's feat was the first in four years; since Suarez achieved it for third time in 18 months</p>

***

## When Goal Difference makes No Difference {-}

Huddersfield's most recent defeat left them in 15th position with a goal difference of -27. Only Stoke(-29) can currently boast(?) a worse figure and one of the current bottom-three, Southampton, are only showing a 15 goal deficit

The correlation between league position and goal difference is clearly going to be high - but there are exceptions

Here is the data for all completed seasons. Hover points for details NB has to be updated to week 38

```{r GD_all_teams_31}

df <- standings %>%
  filter(season < "1995/96" &
           tmYrGameOrder == 42 | season > "1994/95" & tmYrGameOrder == 38) %>%
  select(season, team, cumGF, cumGD, final_Pos) %>%
  arrange(desc(cumGD), desc(cumGF)) %>%
  group_by(season) %>%
  mutate(GDOrder = row_number())

df_2017 <- df %>%
  filter(season == "2017/18")

df_not_2017 <- df %>%
  filter(season != "2017/18")


df %>%
  plot_ly() %>%
  add_markers(
    data = df_not_2017,
    x =  ~ jitter(GDOrder),
    y =  ~ jitter(final_Pos),
    name = "Other Seasons",
    hoverinfo = "text",
    text =  ~ paste(team,
                    "<br>", season,
                    "<br>GD:", GDOrder,
                    "<br>Pos:", final_Pos)
  ) %>%
  add_markers(
    data = df_2017,
    x =  ~ jitter(GDOrder),
    y =  ~ jitter(final_Pos),
    color = I("red"),
    name = "2017/18",
    hoverinfo = "text",
    text =  ~ paste(team,
                    "<br>", season,
                    "<br>GD:", GDOrder,
                    "<br>Pos:", final_Pos)
  ) %>%
  
  layout(
    xaxis = list(title = "Goal Difference Rank"),
    yaxis = list(title = "Final Position")
  ) %>% config(displayModeBar = F, showLink = F)


```
 _N.B. The points have been jittered to avoid overlap_
 
 A couple of outliers include Norwich in the inaugural PL season; they finished 3rd but with a negative goal difference of -4. On the other side of the coin, Birmingham were relegated in 19th place but were only 13th in goal difference
 
 Here are how the relegated teams have fared
 
```{r relegated_31}

df %>%
  arrange(desc(GDOrder)) %>%
  group_by(season) %>%
  slice(1:3) %>%
  plot_ly(
    x =  ~ jitter(GDOrder),
    y =  ~ jitter(final_Pos),
    color = "red",
    hoverinfo = "text",
    text =  ~ paste(team,
                    "<br>", season,
                    "<br>GD:", GDOrder,
                    "<br>Pos:", final_Pos)
  ) %>%
  layout(
    xaxis = list(title = "Goal Difference Rank"),
    yaxis = list(title = "Final Position")
  ) %>% config(displayModeBar = F, showLink = F)


```
 
 Swansea, West Ham and Aston Villa are the teams that would have been relegated if the decision was based on goal difference but actually finished 15th. In Swansea's case, it only delayed the drop by one year
 
--- 
 <p class="factoid">Luka Milivojevic picked up his league leading sixth penalty and is now two ahead of the field in spite of missing a last-minute chance against Man City and having one wrestled off him by Benteke (who missed)</p>
 
---
 



## Tweet of the Week {-}

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Amazingly, although over half of <a href="https://twitter.com/hashtag/premierleague?src=hash&amp;ref_src=twsrc%5Etfw">#premierleague</a> defeats are by a single goal,<br>of  <a href="https://twitter.com/hashtag/htafc?src=hash&amp;ref_src=twsrc%5Etfw">#htafc</a>&#39;s 16 losses this season, 15 have been by 2 or more<br>Only one was v <a href="https://twitter.com/hashtag/MCFC?src=hash&amp;ref_src=twsrc%5Etfw">#MCFC</a> decided by 84th minute goal<a href="https://twitter.com/hashtag/CPFC?src=hash&amp;ref_src=twsrc%5Etfw">#CPFC</a> <a href="https://t.co/MWVCmpEV56">pic.twitter.com/MWVCmpEV56</a></p>&mdash; Andrew Clark (@pssGuy) <a href="https://twitter.com/pssGuy/status/975682892335591426?ref_src=twsrc%5Etfw">March 19, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6732454-thirty-somethings-transferrable-skills-and-tedesco/embed/v4?eid=AQAAAIDVr1qmumYA" title="Audioboom player"></iframe>
***
<p>
 
<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6733025-another-magical-salah-display-and-warnock-s-pub-quiz-football-weekly/embed/v4?eid=AQAAABGgsFrhvGYA" title="Audioboom player"></iframe>

***

## Results and Table {-}


```{r results_31}

weekly_results("2018-03-13","2018-03-20")

```

<p>
---

```{r standings_31}

weekly_table("2018-03-20")

```



## Final Factoid {-}

<p class="factoid">Mane grabs three assists for first time in his EPL career and is just two goals shy of a double-double for a third consecutive season</p>


---



## End of Season Update {-}

This season, only Aguero could match Salah with a hat-trick and an assist in the same game. Prior to them, Kane had rattled the feat off three times in 10 games including the final two matches of last season

---

Further to the original blog post, I have highlighted this year's data but there is nothing exceptional about the campaign

In fact the correlation between position if based on Goal Difference and the actual standings is one of the highest ever

```{r GDu_31}

raw <- df %>%
  ungroup() %>%
  select(-c(cumGF, cumGD, team))

seasons <- df %>%
  select(season) %>%
  unique() %>%
  pull(season)

corr_fun <- function(yr) {
  raw %>%
    filter(season == yr) %>%
    select(-season) %>%
    cor() %>%
    .[1, 2]
}

data <- bind_cols(corr = map_dbl(seasons, corr_fun), season = seasons)

data %>%
  plot_ly(x =  ~ season,
          y =  ~ corr,
          color = ~ season) %>%
  hide_legend() %>%
  layout(
    title = "Correlation between Finishing Position<br>
  and rank based on Goal Difference",
  margin = list(b = 100),
  yaxis = list(title = "Correlation", range = c(0.8, 1)),
  xaxis = list(title = "")
  ) %>% config(displayModeBar = F, showLink = F)
```


---

Milivojevic did finish as top penalty scorer with 7, even after allowing Benteke to take one when they were 4-0 up on Leicester in injury time. He finished three clear of three others (including Murray who missed a couple for Brighton)


---

Mane duly got his double-double, one of just eleven  players to achieve the feat, including team-mates Salah and Firmino.

Even Manchester United have failed to get someone reaching this achievement in 7 of the 26 seasons. Chelsea's 2009/10 had 4 players meet the mark Drogba(29 goals 18 assists); Lampard(22/21); Anelka(11/15); Malouda(12/12)

---

## Code Hints {-}

  base - cor()











