# Week 36 {-}



## Match of the Day {-}

Signs of life from the botton three with wins for WBA and Southampton and a draw for Stoke at Anfield. Unfortunately, all available is the Liverpool game with Italian commentary


```{r topGame_36}

embed_youtube("e7kbmIpByDI", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```

***
<p class="factoid">Somewhat overshadowed by de Bruyne, Sterling (18 goals, 22 Assists) joins Salah as the only other player on the 40 points mark this season</p>

***

## Most wins - win it all {-}

Manchester City have already reached the 30 win mark matching Chelsea's record, set last year, with three games still to play. Manchester United are next with 24

City have already won the title but has there ever been a Premier league campaign
where the most wins has not brought home the trophy? Well the answer is a qualified yes

Here are the podium finishers for every year

```{r most_wins_36}

standings %>%
  filter(res == "Win") %>%
  group_by(season, team, final_Pos) %>%
  tally() %>%
  arrange(desc(n), final_Pos) %>%
  group_by(season) %>%
  slice(1:3) %>%
  arrange(desc(season)) %>%
  select(season, team, position = final_Pos, wins = n) %>%
  DT::datatable(
    width = 400,
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE,
      pageLength = 12,
      lengthMenu = c(6, 12, 18)
    )
  )

```
 
There has been the odd year when two teams have tied for the top spot. Perhaps most notably in 2011/12 when Manchester City scored twice in injury time in the last game of the season to pip Manchester United on goal difference
 
 --- 


## Scoring For and Against {-}

In a no suprise game (Arsenal losing the game in stoppage time), Henrikh Mkhitaryan became the first player to both score for and against Manchester United in the same season

But how often does a player come back to bite the team that sold him

```{r two_club_goals_36}

player_team <- playerGame %>%
  filter(PLAYERID != "OWNGOAL") %>%
  group_by(season, name, TEAMNAME, PLAYERID) %>%
  summarize(tot = sum(Gls)) %>%
  filter(tot > 0)

two_Clubs <- player_team %>%
  ungroup() %>%
  select(-tot) %>%
  group_by(season, name, PLAYERID) %>%
  tally() %>%
  filter(n > 1)

for_against <- function(id, yr) {
  temp <- player_team %>%
    filter(PLAYERID == id & season == yr)
  
  teams <- temp$TEAMNAME
  
  scoredAgainst <- playerGame %>%
    filter(PLAYERID == id & season == yr, Gls > 0) %>%
    pull(Opponents)
  
  team <-
    ifelse(length(intersect(teams, scoredAgainst)) > 0, intersect(teams, scoredAgainst), NA)
  
  data.frame(PLAYERID = id,
             season = yr,
             team = team)
}

data <- map2_df(two_Clubs$PLAYERID, two_Clubs$season, for_against)


goal_ag <- data %>%
  filter(!is.na(team))


player_team %>%
  left_join(goal_ag) %>%
  filter(!is.na(team), team != TEAMNAME) %>%
  arrange(desc(season)) %>%
  select(player = name,
         season,
         `for & against` = team,
         other = TEAMNAME) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

```

Only two players appear twice on the list.

John Hartson performed the feat first with West Ham and Wimbledon the clubs involved

However, this is trumped by Robbie Keane who started the 2002/03 season for Leeds whom he scored against after a transfer to Spurs. Then in 2008/09, he was sold to Liverpool and after a disappointing spell went back to Spurs for 8 million pounds less and... well you can guess the rest


## Best Win {-}

Crystal Palace's 5-0 romp over Leicester both virtually freed them from relegation concerns and was their best ever Premier League win

Here is the best result, in terms of scoreline, for every team that has appeared
 Determing factors are by order, goal difference, goals scored and earliest occurence

```{r big_win_36}

standings %>%
  mutate(GD = GF - GA) %>%
  arrange(desc(GD), desc(GF), gameDate, season) %>%
  group_by(team) %>%
  slice(1) %>%
  select(team, GF, GA, Opponents = OppTeam, date = gameDate) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )


```

Barnsley, Cardiff and Swindon never managed more than a two goal win in their single seasons in the league

***

## Tweet of the Week {-}


```{r}  
shortcode("tweet", "989880967224586240") 
```


## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6834346-good-eggs-girth-and-brains/embed/v4?eid=AQAAAI1M6FqqSGgA" title="Audioboom player"></iframe>
***
<p>
 
<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6834488-boxes-baggies-celtic-sunderland-and-more-football-weekly/embed/v4?eid=AQAAALpM6Fo4SWgA" title="Audioboom player"></iframe>
***

## Results and Table {-}

```{r results_36}

weekly_results("2018-04-24", "2018-05-01")

```

<p>
---

```{r standings_36}

weekly_table("2018-04-24")

```



## Final Factoid {-}

<p class="factoid">Shutout 19 times, three more than any other team, Huddersfield finish with Man City, Chelsea and Arsenal and could yet equal's Derby's 2007/08 EPL record of 22</p>


---

## End of Season Update {-}


#### Most wins - win it all {-}

 How about the converse. Does losing the most games in a season, condemn you to relegation?
 
```{r most_lossesu_36}

standings %>%
  filter(res == "Loss") %>%
  group_by(season, team, final_Pos) %>%
  tally() %>%
  arrange(desc(n), final_Pos) %>%
  group_by(season) %>%
  slice(1:3) %>%
  arrange(desc(season)) %>%
  select(season, team, position = final_Pos, losses = n) %>%
  DT::datatable(
    width = 400,
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE,
      pageLength = 12,
      lengthMenu = c(6, 12, 18)
    )
  )
```
 
Not the same situation at all. In fact, in 1996/97 Spurs finished in the top half of the table in spite of losing almost half their games. With 46 points, they only finished 7 points clear of second-bottom, Middlesbrough 

This season..

With 22 losses you seem pretty well certain to be relegated and even with just 15 a couple of sides have taken the drop including Crystal Palace in a 42 game schedule in 1992/93 and ..


#### Best Win {-}

On the same theme of opposites, here are the worst losses suffered by each team

The criteris this time is biggest goal difference; lowest goals for; and game date

```{r worst_lossU_36}


standings %>%
  mutate(GD = GF - GA) %>%
  arrange(GD, GF, gameDate, season) %>%
  group_by(team) %>%
  slice(1) %>%
  select(team, GF, GA, Opponents = OppTeam, date = gameDate) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

```

Unsurprisngly, Manchester United lead the way as putting most teams to the sword, eight, although none for more than seven years

Other than the teams appearing in the Premier League for the first time, Brighton and Huddersfield, Watford - and more surprisngly, Liverpool suffered their worst defeats. In mitigation , against Manchester City, the Merseyside team had Mane sent off after 37 minutes , won the return fixture and also beat their opponents home and away in the Champions League

--

## Code Hints

  * base - intersect()











