# Week 38 {-}


Very little to play for - other than stacks of place money
Well at least we got a nine goal thriller. And Lamela scored twice, his first of the season!


***

<p class="factoid">Lucas Moura records the 50th 4-assist game in EPL history on the final day of the season, matching Spurs teammate Eriksen who did the same trick on the last day of 2016/17</p>

***

## Man City Records {-}

City set a plethora of records including for wins, points and goals in an EPL season. They were also only the second team (Man. Utd. 1999/00) to have six players reach the 20 Goals and Assists combined mark

They also very rarely trailed in games


```{r mins_behind_38}


data <- scoreLines %>%
  left_join(teamGames) %>%
  left_join(standings, by = c("MATCHID" = "MATCHID", "TEAMNAME" = "team")) %>%
  select(
    team = TEAMNAME,
    season = season.x,
    gameDate = gameDate.x,
    mins,
    diff,
    tmYrGameOrder = tmYrGameOrder.x,
    GF = GF.y,
    GA = GA.y,
    OppTeam,
    final_Pos
  )


test <- data %>%
  rename(TEAMMATCHID = TEAMMATCHID.x) %>%
  filter(season == "2017/18" & team == "Man. City") %>%
  mutate(
    up = if_else(diff > 0, 1, 0),
    down = if_else(diff < 0, 1, 0),
    even = if_else(diff == 0, 1, 0)
  ) %>%
  group_by(TEAMMATCHID, tmYrGameOrder) %>%
  summarize(up = sum(up),
            down = sum(down),
            even = sum(even)) %>%
  gather(category, time, -c(tmYrGameOrder, TEAMMATCHID))


## team by game - but overall not by minute
test %>%
  plot_ly(x =  ~ tmYrGameOrder, y =  ~ time) %>%
  add_bars(color =  ~ category) %>%
  layout(
    barmode = "stack",
    yaxis = list(title = "Mins"),
    xaxis = list(title = "")
  )


```

and the total of 156  minutes total time in arrears is the lowest in Premier League history

---

Here is every team's record for time up/down/even over a full season


```{r up_down_38}




data %>%
  ungroup() %>%
  group_by(season, team) %>%
  mutate(
    up = if_else(diff > 0, 1, 0),
    down = if_else(diff < 0, 1, 0),
    even = if_else(diff == 0, 1, 0)
  )  %>%
  summarize(up = sum(up),
            even = sum(even),
            down = sum(down)) %>%
  arrange(down) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

```


---

<p class="factoid">In spite of their dismal start, Palace scored at least two goals at home on eleven occasions, bettering teams like Spurs(9) and Chelsea(8)</p>

---



## Leaving with a Bang {-}

Well Swansea certainly didn't as they went from appearing safe - with a fall-back position of the final two home games against Southampton and Stoke - to the drop

On the other hand, Stoke and, to a greater extent, WBA at least made it interesting. Here are the number of losses relegated teams (since the 38 season began) have suffered in their final five games


```{r down_38}


relegated <- standings %>%
  filter(season > "1994/95") %>%
  select(season, team, final_Pos) %>%
  unique() %>%
  group_by(season) %>%
  arrange(desc(final_Pos)) %>%
  slice(1:3) %>%
  select(season, team)

losses_last_games <- function(x, y)  {
  standings %>%
    filter(team == x,
           season == y,
           between(tmYrGameOrder, 34, 38),
           res == "Loss") %>%
    tally()
}



losses <-
  map2_df(relegated$team, relegated$season, losses_last_games)

df <- bind_cols(relegated, losses)

df %>%
  arrange(n, desc(season)) %>%
  rename(losses = n) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )


```

 Swansea leave the division on the back of five consecutive defeats, a rarity even for the poorest sides
 
 It is the only season in which two teams only lost once at the end. Both
 Stoke and WBA suffered at the hands of Crustal Palace

***

## Tweet of the Week {-}

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr"><a href="https://twitter.com/hashtag/AFC?src=hash&amp;ref_src=twsrc%5Etfw">#AFC</a> no away points after 7 <a href="https://twitter.com/hashtag/EPL?src=hash&amp;ref_src=twsrc%5Etfw">#EPL</a> games in 2018<br>Previous years range from 6 -19 (2002)<a href="https://twitter.com/ArsenalFanTV?ref_src=twsrc%5Etfw">@ArsenalFanTV</a> <a href="https://t.co/Xa8VzG5N44">pic.twitter.com/Xa8VzG5N44</a></p>&mdash; Andrew Clark (@pssGuy) <a href="https://twitter.com/pssGuy/status/994375363219238912?ref_src=twsrc%5Etfw">May 10, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


## Totally Football and Guardian Weekly Review Podcasts {-}

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6854834-bad-lieutenants-ugly-town-and-extra-swazz/embed/v4?eid=AQAAAG0S-lqymGgA" title="Audioboom player"></iframe>
***
<p>
 
<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6855057-wenger-gone-goal-frenzies-and-tranmere-rovers-return-football-weekly/embed/v4?eid=AQAAALwS-lqRmWgA" title="Audioboom player"></iframe>
***

## Results and Table {-}

```{r results_38}

weekly_results("2018-05-07", "2018-05-14")

```

<p>
---

```{r standings_38}

weekly_table("2018-05-14")

```

## Final Factoid {-}

<p class="factoid">Burnley's, Dwight McNeil was the only debutant on the last day - the fewest since there were none in 2011/12. He came on in the 89th minute and Bournemouth scored the winner a minute later</p>


---

## End of Season Update {-}

It's the end of the season. There aren't any! 





