# Week 25 {-}


## Match of the Day {-}


Spurs score early - but should have more often - against a lack-lustre Manchester United sporting their shiny new toy, Alexis Sanchez 



```{r topGame_25}

embed_youtube("twxvaC-gP2E", width = 420, height = 315, frameborder = 0,
   allowfullscreen = TRUE, query = NULL) 


```


***

## One that got away, not {-}

Riyad Mahrez stayed at Leicester, in spite of a transfer request and rumours that Manchester City were willing to spend somewhere around 75 million on him

There is no doubt that he has rediscovered the form he showed in the title-winning season. At the time of writing he is contributing to over 60% of his team's goals

```{r mahrez_25 }

tmGoals <- standings %>%
  group_by(team) %>%
  filter(season == "2017/18" & tmGameOrder == max(tmGameOrder)) %>%
  select(team, cumGF)

playerGame %>%
  group_by(name, PLAYERID, TEAMNAME) %>%
  filter(season == "2017/18") %>%
  summarize(points = sum(Gls) + sum(Assists)) %>%
  rename(team = TEAMNAME) %>%
  left_join(tmGoals) %>%
  filter(points > 0) %>%
  mutate(pc = round(100 * points / cumGF, 1)) %>%
  arrange(desc(pc)) %>%
  ungroup() %>%
  select(name, team, `%` = pc) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = FALSE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )

```

_N.B. Up to 2 assists may be credited to each goal_

On publishing, the under-the-radar leader is Brighton's Pascal Gross. Albeit helped by the team's poor scoring record, he has proven a shrewd purchase




***

<p class="factoid">Chelsea's surprising capitulation to Bournemouth means that they have now conceded 943 goals in the Premier League, tied with Man. City and two less than Arsenal. Spurs lead with 1252</p>
***

## Goal machines {-}

Both Liverpool and Manchestr City recorded 3-0 victories mid-week and are tied for  winning games by three goal margins,at 9

City just shade Liverpool 14-13 in the number of occasions they have scored at least three goals

Here are the leaders for each season at the 25 game mark

```{r 3GF_25}

standings %>%
  filter(GF >= 3 & tmYrGameOrder <= 25) %>%
  group_by(team, season) %>%
  tally() %>%
  group_by(season) %>%
  arrange(desc(n)) %>%
  slice(1) %>%
  plot_ly(x =  ~ season,
          y =  ~ n,
          color =  ~ team) %>%
  add_bars() %>%
  layout(
    title = "Team with most 3+ goal games after 25 Rounds",
    xaxis = list(title = ""),
    margin = list(b = 80),
    yaxis = list(title = "Count")
  ) %>%  config(displayModeBar = F, showLink = F)

         
```

***


## Totally Football and Guardian Weekly Review Podcasts

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6633122-swoops-swaps-and-resurgent-swans/embed/v4?eid=AQAAALr1dFqiNmUA" title="Audioboom player"></iframe>
***
<p>

<iframe width="100%" height="300" style="background-color:transparent; display:block; padding: 0; max-width: 700px;" frameborder="0" allowtransparency="allowtransparency" scrolling="no" src="//embeds.audioboom.com/posts/6633374-spurs-sweep-manchester-united-aside-plus-deadline-day-delights-football-weekly-extra/embed/v4?eid=AQAAAFX2dFqeN2UA" title="Audioboom player"></iframe>

***

## Results and Table

```{r results_25}

weekly_results("2018-01-28","2018-02-02")

```

---

```{r standings_25}

weekly_table("2018-02-02")

```


## Final Factoid

<p class="factoid">Sixth time Man. Utd. have conceded in first minute (Everton lead with 10) and second time to Spurs. Eriksen joining Jermain Defoe's effort in 2009/10</p>


---

## End of Season Update {-}



#### One that got away, not

Although slipping somewhat, Mahrez still led Leicester in contributing to goals scored
but the real star of the league, and still somewhat appreciated, was Pascal Gross

Obviously, with a low scoring team you are going to have a chance to shine more but Brighton's bargain buy was exceptional. See how he stacks up all-time

```{r mahrezu_25}

##looks like still might take too long hangs when building book

df <- playerGame %>%
  group_by(name, PLAYERID, TEAMNAME, season) %>%
  summarize(points = sum(Gls) + sum(Assists)) %>%
  rename(team = TEAMNAME) %>%
  left_join(tmGoals) %>%
  filter(points > 19) %>%
  mutate(pc = round(100 * points / cumGF, 1)) #482

df_2017 <- df %>%
  filter(season == "2017/18")


df %>%
  plot_ly(x =  ~ points,
          y =  ~ pc,
          color = I("lightgrey")) %>% # 10000 points if points>0
  add_markers(
    hoverinfo = "text",
    name = "Previous Seasons",
    text =  ~ paste0(
      name,
      "<br>",
      team,
      "<br>",
      season,
      "<br>Pts: ",
      points,
      "<br> %: ",
      pc
    )
  ) %>%
  add_markers(
    data = df_2017,
    x =  ~ points,
    y =  ~ pc,
    color = I("red"),
    name = "2017/18",
    hoverinfo = "text",
    text =  ~ paste0(
      name,
      "<br>",
      team,
      "<br>",
      season,
      "<br>Pts: ",
      points,
      "<br> %: ",
      pc
    )
  ) %>%
  layout(
    title = "Player Goals and Assists (min 20)<br> as % of Team Goals",
    xaxis = list(title = "Goals + Assists"),
    yaxis = list(title = "Proprtion of Team Goals")
  ) %>% config(displayModeBar = F, showLink = F)

```

In order to avoid a big delay in processing the minimum points total for season is 20. Just under 500 players have achieved tha in 26 seasons

---


```{r GAU_25}

standings %>%
  group_by(team) %>%
  summarize(
    totGls = sum(GA),
    seasons = floor(n() / 38),
    av = round(totGls / seasons, 0)
  ) %>%
  arrange(desc(totGls)) %>%
  select(team, conceded = totGls, seasons, av) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    rownames = TRUE,
    options = list(
      paging = TRUE,
      searching = TRUE,
      info = FALSE
    )
  )
```

---

Man City led the way in games scoring at least 3 goals .. just shy of Liverpools 2013/14 record.. mark 21

---

## Code Hints {-}

  * none this week





