--- 
title: "Premier League Annual 2017/18"
author: "Andrew Clark"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
#bibliography: [book.bib, packages.bib] 
#biblio-style: apalike
link-citations: yes
description: "This is comprised of a  compilation of weekly
blog posts, extensively extended with end-of-season updates, featuring varying aspects of the EPL using R code
to produce charts and tables. Tweets, pods and game videos are also embedded."
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=TRUE,  fig.align = 'center')
```

```{r setup}


## Required Libraries

library(tidyverse) #for data carpentry
library(plotly) # for plots
library(DT) # for tables
library(vembedr) # for videos
library(blogdown) # for shortcode(tweet)
library(crosstalk)
#library(forcats)
library(glue)
library(htmltools)
library(RcppRoll)
library(sparkline)
library(lubridate)
library(grid)
library(VennDiagram)
library(explodingboxplotR)
library(data.table)


## Required data files

standings<- readRDS("data/soccer/standings.rds")
hth<- readRDS("data/soccer/hth.rds")
managerGame<- readRDS("data/soccer/managerGame.rds")
managers<- readRDS("data/soccer/managers.rds")
matchTeam<- readRDS("data/soccer/matchTeam.rds")
goals<- readRDS("data/soccer/goals.rds")
playerGame<- readRDS("data/soccer/playerGame.rds")
playerClub<- readRDS("data/soccer/playerClub.rds")
teamGames<- readRDS("data/soccer/teamGames.rds")
#teamCodes<- readRDS("data/soccer/teamCodes.rds") Occasional issues
assists<- readRDS("data/soccer/assists.rds")
scoreLines<- readRDS("data/soccer/scoreLines.rds") #
scores<- readRDS("data/soccer/scores.rds")
summary<- readRDS("data/soccer/summary.rds")
Place <- readRDS("data/soccer/Place.rds")
Play  <- readRDS("data/soccer/Play.rds")
Method <- readRDS("data/soccer/Method.rds")

teamCodes <- teamGames %>%
  ungroup() %>%
  select(TEAMNAME,TEAMID) %>%
  unique()

managers[is.na(managers$Left),"Left"] <- as.Date(Sys.Date(), origin= '1970-01-01') #

managerGame <-managers %>% 
  mutate(name=paste(FirstName,Lastname)) %>% 
  group_by(ManagerID,ManagerTeam) %>% 
  inner_join(standings,by=c("TEAMNAME"="team")) %>% 
  select(Lastname,FirstName,name,ManagerID,ManagerTeam,Joined,Left,TEAMNAME,gameDate,res,GF,GA,position) %>% 
  filter(gameDate>=as.Date(Joined)&gameDate<=as.Date(Left)) %>% 
  mutate(points=ifelse(res=="Win",3,ifelse(res=="Draw",1,0))) %>% 
  ungroup()

## Functions for weekly reslts and tables

weekly_results <- function(startDate,endDate) ({
  
hth %>% 
  filter(season=="2017/18"&gameDate>=startDate&gameDate<=endDate) %>% 
  filter(venue=="H") %>% 
  arrange(team) %>% 
  select(Home=team,GF,GA,Away=OppTeam) %>% 
  DT::datatable(rownames=FALSE,class='compact stripe hover row-border',colnames = c('', '', '', ''),
                options= list(paging = FALSE, searching = FALSE,info=FALSE,
                              columnDefs = list(list(width = '40%', targets = list(0,3)))),width=350)
  
}) 

weekly_table <- function(endDate) ({
  
  hth %>%
  filter(season == "2017/18" & gameDate <= endDate) %>%
  group_by(team) %>%
  mutate(
    W = ifelse(res == "Win", 1, 0),
    L = ifelse(res == "Loss", 1, 0),
    D = ifelse(res == "Draw", 1, 0)
  ) %>%
  summarise(
    P = n(),
    Pts = sum(points),
    W = sum(W),
    D = sum(D),
    L = sum(L),
    GD = sum(GF) - sum(GA),
    GF = sum(GF)
  ) %>%
  arrange(desc(Pts), desc(GD), desc(GF), team) %>%
  DT::datatable(
    class = 'compact stripe hover row-border order-column',
    colnames = c('', 'P', 'Pts', 'W', 'D', 'L', 'GD', 'GF'),
    rownames = TRUE,
    options = list(
      paging = FALSE,
      searching = FALSE,
      info = FALSE
    )
  )
  
  
}) # shows as error but cannot see where and seems to work must be something above

### function extracted from doBy package which takes too long to load
subSeq <- function (x, item = NULL) {
  rrr <- rle(x)
  len <- rrr$lengths
  val <- rrr$values
  
  first <- last <- rep.int(NA, length(val))
  first[1] <- 1
  last [1] <- len[1]
  if (length(val)>1){
    for (kk in 2:length(val)){
      first[kk] <- last[kk-1]+1
      last [kk] <- last[kk-1]+len[kk]
    }
  }
  midp <- floor(first + len/2)
  
  ans <- cbind(first=first, last=last, slength=len, midpoint=midp)
  
  if (!is.null(item)) {
    iii <- val==item
    ans <- as.data.frame.matrix(ans[iii,,drop=FALSE], stringsAsFactors=FALSE)
    ans$value <- val[iii]
  } else {
    ans <- as.data.frame.matrix(ans, stringsAsFactors=FALSE)
    ans$value <- val
  }
  ans
}

```

# Introduction {-}

Welcome to my - and possibly a world - first interactive sports annual

The basis is the weekly [blog posts]() I have produced over the course of the 2017/18 English Premier League. It has been expanded with some additional - and I have passed commentary in many of the chapters updating the analyses reached at that stage of the season

All the code is in R and, for those of you inclined, you can observe the code. I am no expert
so if you have any hints to .. you can . The whole and the packages and code I used. Some weeks I will list some packages and functions which I may have used for the first time and may prove helpful

The data ...

Charts are generally zoomable/panable and often more detailed information can be found by hovering markers

Help wanted
-----------

* Enhance code
* Enhance design
* Translation foreign languages
* Other seasons
