---
title: "blog Code WD"
output: html_notebook
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message=FALSE, echo=FALSE)
```


```{r startup}

library(tidyverse)
library(plotly)
library(odbc)
library(DT)
library(glue)
library(htmltools)


con <- dbConnect(odbc::odbc(),
                   Driver    = "SQL Server", 
                   Server    = "sqlb12.webcontrolcenter.com",
                   Database  = "epldb",
                   UID       = "soccer",
                   PWD       = "mlcepl",
                   Port      = 1433)



```


## some issues so just use rds/featger files to start with

```{r weekly_update}

#dbWriteTable(con, "managerTeam", managerTeam) #??
dbWriteTable(con, "allPlayers", allPlayers, overwrite = TRUE)
#dbListTables(con) %>% head()
dbWriteTable(con, "playerGame", playerGame, overwrite = TRUE)
dbWriteTable(con, "playerClub", playerClub) #Cannot find data type DATE.
dbWriteTable(con, "summary", summary, overwrite = TRUE)
dbWriteTable(con, "standings", standings, overwrite = TRUE)
dbWriteTable(con, "leaders", leaders, overwrite = TRUE)
dbWriteTable(con, "allGoalsPlayer", allGoalsPlayer, overwrite = TRUE)
dbWriteTable(con, "goals", goals, overwrite = TRUE)
dbWriteTable(con, "Goals_team_for", Goals_team_for, overwrite = TRUE)
dbWriteTable(con, "Goals_team_ag", Goals_team_ag, overwrite = TRUE)
dbWriteTable(con, "Play", Play, overwrite = TRUE)
dbWriteTable(con, "Method", Method, overwrite = TRUE)
dbWriteTable(con, "Place", Place, overwrite = TRUE)
dbWriteTable(con, "teamGames", teamGames, overwrite = TRUE)
dbWriteTable(con, "milestones", milestones, overwrite = TRUE)
dbWriteTable(con, "managers", managers, overwrite = TRUE)
dbWriteTable(con, "goalSeqs", goalSeqs, overwrite = TRUE)
dbWriteTable(con, "goalSeqsClub", goalSeqsClub, overwrite = TRUE)
dbWriteTable(con, "hth", hth, overwrite = TRUE)
dbWriteTable(con, "goalSeqTeam", goalSeqTeam, overwrite = TRUE)

```



# Round 2

 * crowds
 * 4-0 wins
 * goals without reply (inc last season-start season)
 * strength of schedule
 * not scoring in 
 * saints end barren home run
 



```{r}

```



# Round 1

## Foreign Players

One fairly constant complaint has been the downturn in English player over the years. T

```{r}


games <- teamGames %>% 
  ungroup() %>% 
  filter(tmYrGameOrder==1) %>% 
  pull(MATCHID)

 playerGame %>% 
  filter(MATCHID %in% games&START>0) %>% 
  group_by(season,COUNTRY) %>% 
  summarise(tot=n()) %>% 
  group_by(season) %>% 
  mutate(pc=round(100*tot/sum(tot),1)) %>% 
  filter(COUNTRY=="England") %>% 
  ungroup() %>% 
  plot_ly(x=~season,y=~pc) %>% 
  add_lines() %>% 
  layout(margin=list(b=70),
         title="% of Starters born in England - EPL Round 1",
    xaxis=list(title=""),
    yaxis=list(title="% ")
  )

```

This appears to have stabilized over past coupld of seasons

```{r}

# 1st weeek tis year
playerGame %>% 
  filter(MATCHID %in% games&START>0&season=="2017/18") %>% 
  #mutate(place=glue({city},{COUNTRY})) %>% 
  mutate(place=paste0(city,", ",COUNTRY)) %>% 
  select(team=TEAMNAME,player=LASTNAME,place) %>% 
  arrange(team,player)%>%
                         DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE,lengthMenu=c(11,11)))


playerGame %>% 
  filter(MATCHID %in% games&START>0&season=="2017/18") %>% 
  group_by(TEAMNAME,COUNTRY) %>% 
  tally() %>% 
  ungroup() %>% 
  select(TEAMNAME) %>% 
  group_by(TEAMNAME) %>% 
  tally()


```

New-signing, Lacazette, is the only Frenchan on Arsenals week 1 team. Stoke boast ten different nationalities

 

```{r }

country.df <- playerGame %>% 
  filter(MATCHID %in% games&START>0&season=="2017/18") %>% 
  group_by(COUNTRY) %>% 
  tally() 

# works but need to change country ideally to alphabetical -and rev doe not seem to hel
 country.df$COUNTRY <-  factor(country.df$COUNTRY, levels = country.df$COUNTRY[order(country.df$n,country.df$COUNTRY)])
 
  
country.df  %>% 
  plot_ly(x=~n,y=~COUNTRY,height=1000) %>% 
  layout(margin=list(l=100))

```
 
No clear candidate for second place but newly-promoted Huddersfield coach, David Wagner, has signed a few fellow-Germans
Only two Scots, Fraser (Bournemouth) and 33-year-old, Fletcher (Stoke). The last time Scotland held the second rank after England in a lead-off game was in 1999


```{r 2nd}


 # playerGame %>% 
 #  filter(MATCHID %in% games&START>0) %>% 
 #  group_by(season,COUNTRY) %>% 
 #  summarise(tot=n()) %>% 
 #  group_by(season) %>% 
 #  arrange(desc(tot)) %>% 
 #  slice(2)

```


## Big Away Wins


```{r, fig.height= 12}

# first games for teams that have ghad 3 goals margins
didIt <-standings %>% 
  arrange(gameDate) %>% 
  filter(venue=="A") %>% 
  group_by(team) %>% 
  mutate(awayGameOrder=row_number()) %>% 
    filter((GF-GA)>=3) %>% 
  group_by(team) %>% 
    slice(1)

# teams that have never achieved  
never <-setdiff(unique(standings$team),unique(didIt$team))

# never teams away games to date - select relevant cariables+ add ever
neverTeams <- standings %>% 
  filter(team %in% never&venue=="A") %>% 
  group_by(team) %>% 
  arrange(tmGameOrder) %>% 
  mutate(awayGameOrder=row_number()) %>% 
  arrange(desc(awayGameOrder)) %>% 
  slice(1) %>% 
  select(team,awayGameOrder) %>% 
  mutate(ever="Never")

# cf above re fields
didItTeams <- didIt %>% 
  select(team,awayGameOrder) %>% 
  mutate(ever="First")

# combine
df <- rbind(neverTeams,didItTeams)

## order team as factor for chart
df$team <-  factor(df$team, levels = df$team[order(df$awayGameOrder)])


 #chart 
df %>% 
  plot_ly(x=~awayGameOrder,y=~team,height=1000) %>% 
  add_bars(color=~ever) %>% 
  layout(title=" EPL Away Games before 3 goal victory",
         margin=list(l=120),
         xaxis=list(title="Games"),
         yaxis=list(title="")
         
         ) %>%  config(displayModeBar = F,showLink = F)


```

Somewhat surprising that four teams, including some unfancied ones started their Premier League away encounters in such a splendid fashion

  Southampton definitely buck the trend taking 186 games to reach the target - though they have had more success recently
  
```{r}
  
  standings %>% 
      filter(venue=="A"&team=="Southampton") %>% 
    mutate(GD=GF-GA) %>% 
    group_by(GD) %>% 
    tally() %>% 
    plot_ly(x=~GD,y=~n) %>% 
    add_bars(color=I("red"))
  
  
  
```


```{r}

sakho <-playerGame %>% 
  filter(PLAYERID=="SAKHOM"&TEAMNAME=="Crystal P"&mins==90) %>% 
  pull(MATCHID)

standings %>% 
  filter(MATCHID %in% sakho&team=="Crystal P") %>% 
  select(res) #5/1/1

standings %>% 
  filter(team=="Crystal P") %>% 
  arrange(desc(gameDate)) %>% 
  mutate(win=ifelse(res=="Win",1,0)) %>% 
  mutate(cumWins=cumsum(win)) %>% 
  select(gameDate,res,cumWins)

```


## Man Utd are back

I know, only West Ham, but still far from the turgid play of recent seasons


```{r}



```



<p class="factoid">Crystal Palace have won just nine of their last 33 league games. More than half these victories came in  the 7 games in which Sakho played a full game</p> 




Man U winning by 4 - copy twitter

surprise of week - Chelsea 



 





```
week 2 want to exyend to any player better
Try it out for any player

```{r}

pl <-playerGame %>% 
  group_by(PLAYERID) %>% 
  select(PLAYERID,name) %>% 
  unique()

players <- pl$PLAYERID
names(players) <-pl$name 
  


sd <- SharedData$new(data)

fs <- filter_select(
id = "playerID", #id	An HTML element ID; must be unique within the web page
label = "Select Player",
sharedData = sd,
group =  ~ players, 
allLevels = FALSE,
multiple = FALSE
)


## this is needed as crosstalk does not work nicely with bootstrap, apparently
fs_nobootstrap <- fs

attr(fs_nobootstrap, "html_dependencies") <- Filter(
  function(dep) {dep$name != "bootstrap"},
  attr(fs_nobootstrap, "html_dependencies")
)

tagList(
  fs_nobootstrap,
  br()
)

```
