---
title: "Daff Example"
author: "pssguy"
date: "August 18, 2015"
output: html_document
---

This is a very quick look at the [daff package](https://github.com/paulfitz/daff)

```{r, warning=FALSE, message=FALSE}
# Main libraries
library(daff)
library(tidyr)


library(DT)
library(readr)
library(dplyr)
```
The package is used to show the difference between two data.frames   

I will construct them with just one value differing, which is what I wish to have as my final output

```{r}

df_a <- data.frame(x=c("a","b"), y=c(1,2))
df_b <- data.frame(x=c("a","b"), y=c(1,3))

```
The function data_diff() shows the differences in the form of a TableView.
  As I understand (probably incorrectly) to get it into data.frame format you need to write it to a csv fie and then retrieve

```{r}
dd <- diff_data(df_a,df_b)
write_diff(dd, "diff.csv")
res <- read_csv("diff.csv")
print(res)
```

I now want to separate out the y column, only showing rows where changes have occurred
and restrict the output to the relevant columns. I can easily do this using the tidyr,DT and ddplyr packages

```{r}

separate(data=res,col= y,into=c("Old","New"),sep="->", extra="drop") %>% 
      filter(!is.na(New)) %>% 
      select(x,New) %>% 
  DT::datatable(rownames=FALSE,width=200,class='compact stripe hover row-border',options= list(paging = FALSE, searching = FALSE,info=FALSE, columnDefs = list(list(className = 'dt-right', targets = 1))))

```
I will be using this shortly to show weekly milestones achieved in the Barclays Premier League
For updates follow me on [twitter](https://twitter.com/pssGuy)
