---
title: "SA maps working doc"
output: html_notebook
---

Working doc for poss sa speech
would be nice to be able to extend elsewhere/use hexmapr

The data comes from [maplibrary.org](http://maplibrary.org/library/stacks/Africa/South%20Africa/index.htm) so has lots of different maps from around world. I downloaded shp files but there asre other formats dra/ etc.
Not obvious there is any sf maps available yet

There are 9 subdivison areas in south africa - which also have downloadable data (what is white gap)

[robin lovelace book](https://bookdown.org/robinlovelace/geocompr/spatial-class.html)

not 

http://www.diva-gis.org/datadown looks like may also have info (could be just maps/pop - which are in grids)

```{r}

# Load libraries

# General usage
library(tidyverse) #  data-carpentry
library(plotly) # interactive graphics
library(DT) #  tables
library(leaflet) #  maps

# Topic specific
library(sp)
library(sf)

```

```{r}
## read in shape file
library(maptools)

# unzipped the SOU_admin_SHP.zip file gives SOU.dbf/SOU.shp/SOU.shx files in the shapefiles directory

data <- readShapePoly("shapefiles/SOU.shp")
#use rgdal::readOGR or sf::st_readError in getinfo.shape(filen) : Error opening SHP file


data <- st_read("shapefiles/SOU.shp", stringsAsFactors = FALSE)
# Error in CPL_read_ogr(dsn, layer, as.character(options), quiet, type,  : 
#   Open failed. also same with /shapefiles/SOU.shp 

# from maptools vignette
library(maptools)
xx <- readShapePoly(system.file("shapes/sids.shp", package="maptools")[1],
 IDvar="FIPSNO", proj4string=CRS("+proj=longlat +ellps=clrk66"))  #SPDF
plot(xx, border="blue", axes=TRUE, las=1)
#plot.new() not in vignette
text(coordinates(xx), labels=row.names(xx), cex=0.6) #Error in text.default(coordinates(xx), labels = row.names(xx), cex = 0.6) : plot.new has not been called yet
as(xx, "data.frame")[1:5, 1:6]
xxx <- xx[xx$SID74 < 2,]
plot(xxx, border="red", add=TRUE)
tmpfl <- paste(tempdir(), "xxpoly", sep="/")
writePolyShape(xxx, tmpfl)
getinfo.shape(paste(tmpfl, ".shp", sep=""))
axx <- readShapePoly(tmpfl, proj4string=CRS("+proj=longlat +ellps=clrk66"))
plot(xxx, border="black", lwd=4)
plot(axx, border="yellow", lwd=1, add=TRUE)
unlink(paste(tmpfl, ".*", sep=""))


# https://edzer.github.io/UseR2017/#sf-handling-real-data good sf recource

fname = system.file("shape/nc.shp", package = "sf")
nc = read_sf(fname)

gname="C:/Users/Andrew/Documents/R/southAfrica/shapefiles/SOU.shp"
nc2 = read_sf(gname)
# Error in CPL_read_ogr(dsn, layer, as.character(options), quiet, type,  : 
#   NULL error in sfc_from_ogr addressed https://github.com/r-spatial/sf/issues/351

#library(rgdal) # prev loaded via namespace so did not load

data <- rgdal::readOGR(gname)

# Deleted feature IDs:  334OGR data source with driver: ESRI Shapefile 
# Source: "C:/Users/Andrew/Documents/R/southAfrica/shapefiles/SOU.shp", layer: "SOU"
# with 334 features
# It has 18 fields

class(data)
# > class(data)
# [1] "SpatialPolygonsDataFrame"
# attr(,"package")
# [1] "sp"

plot(data)

# world_sp = as(world, Class = "Spatial")
data_sf = st_as_sf(data, "sf")
class(data_sf) #[1] "sf"         "data.frame"
names(data_sf)

#  [1] "ID"         "LBL"        "FIP"        "MMT_ID"     "SHORT__FRM" "LONG_FRM"   "ADM0"       "ADM1"       "ADM2"       "ADM3"      
# [11] "ADM4"       "ADM5"       "STL.0"      "STL.1"      "STL.2"      "STL.3"      "STL.4"      "STL.5"      "geometry"  


```

No data yet doemlodaed as csv from data portal e.g http://southafrica.opendataforafrica.org/cuyaur/household-expenditures-and-living-conditions-of-south-africa-2009
Next release date: Tuesday, 19 December 2017


```{r}

hhExpenditure <- read_csv("data/hhExpenditure.csv")
names(hhExpenditure)

unique(hhExpenditure$location)

#[1] "location" "variable" "sex"      "Unit"     "Date"     "Value" 

```


 admin areas are by th regions 
 
```{r}

region_sp <- rgdal::readOGR("C:/Users/Andrew/Documents/R/southAfrica/shapefiles/SOU-level_1.shp")
region_sf= st_as_sf(region_sp, "sf") #22 x5 but seems to break down west cape
plot(region_sf) # as a test

## try splitting off the other rows

region_sf <- region_sf %>% 
  filter(!is.na(CAPTION))

plot(region_sf)

```
 
```{r}

# now look at data


food <- hhExpenditure %>% 
filter(str_sub(variable,1,4)=="Food")

region_sf <- region_sf %>% 
  mutate(location=as.character(CAPTION))

df <- food %>% 
  left_join(region_sf,by=c("location"="CAPTION"))

food$location
region_sf$CAPTION

glimpse(food)
glimpse(region_sf)

#  geom_text(aes(V1, V2, label = substr(name,1,4)), size=4,color = "white") form hexMapCreation.Rmd
ggplot(df) +
  geom_sf(aes(fill = Value)) +
  geom_text(aes())
```

ggplot vignette

```{r, warning = FALSE}

#https://bhaskarvk.github.io/user2017.geodataviz/presentations/06-Static_Maps.html#12

library(sf)
world <- st_as_sf(rnaturalearth::countries110)
europe <- dplyr::filter(world, region_un=="Europe" & name!='Russia')
# A bounding box for continental Europe.
europe.bbox <- st_polygon(list(
  matrix(c(-25,29,45,29,45,75,-25,75,-25,29),byrow = T,ncol = 2)))
europe.clipped <- st_intersection(europe, st_sfc(europe.bbox, crs=st_crs(europe)))
ggplot(europe.clipped, aes(fill=gdp_md_est/1000)) +
  geom_sf(alpha=0.8,col='white') +
   coord_sf(crs="+proj=aea +lat_1=36.333333333333336 +lat_2=65.66666666666667 +lon_0=14") +
  hrbrthemes::theme_ipsum_rc() +
  viridis::scale_fill_viridis(
    name='Median GDP \n(in Billions)', direction = -1, labels=scales::dollar) +
  labs(x=NULL, y=NULL, title="Some Title",
       caption='Source: http://www.naturalearthdata.com/')
```

ot tmap

```{r}

library(tmap)
current.mode <- tmap_mode("plot")

data(World, Europe, metro)

tm_shape(World) +
    tm_text("name", size="AREA")

tm_shape(Europe) +
    tm_polygons() +
    tm_text("iso_a3", size="AREA", col = "grey20", root=4, shadow = TRUE, scale=2,
        size.lowerbound = .1) +
tm_shape(Europe) +
    tm_text("name", size="AREA", root=4, scale=1,
        ymod=-1 * tmaptools::approx_areas(Europe, target = "norm")^(1/4))

## in view mode automatically goes to leaflet
tm_shape(df) +
    tm_polygons() #Error in get_projection(mshp_raw, as.CRS = TRUE) : shp is neither a Spatial nor a Raster object need to add a CRS


```

think about colorblindr package

```


```
 
 Has other household data e.g possession of 