---
title: "R Notebook"
output: html_notebook
---

extending what i did in blog witj remoteDatabases and further in draft dbConnections.Rmd

That created databases as data.frames first I want to go away from tht
I now port the processed files from UpdatingSQL so can use directly in code

Can then revalp github which should work - make package or whatever


```{r, message=FALSE, warning=FALSE}


library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(plotly)
library(knitr)
#ggplot2_2.2.1.9000 also gets attached (whihc is useful as it is used in the fortify) # alt could load ggfortify
library(stringr)

```


```{r, echo=FALSE}

con <- dbConnect(odbc::odbc(),
                   Driver    = "SQL Server", 
                   Server    = "sqlb12.webcontrolcenter.com",
                   Database  = "epldb",
                   UID       = "soccer",
                   PWD       = "mlcepl",
                   Port      = 1433)

```

check connection
```{r}
dbListFields(con, "playerGame")

playerGame <-tbl(con, "playerGame") # still alist
# playerGame <- playerGame %>% 
#   as.data.frame()

#pg <- playerGame %>% fortify() # as with as.data.frame takes an age - may be best done in individual code as below


system.time(
playerGame <-tbl(con, "playerGame") %>% 
  select(PLAYERID,name) %>% 
   collect() %>% 
   unique()
) 
  # user  system elapsed 
  #  2.55    0.17   41.94 


system.time(
gb <-tbl(con, "playerGame") %>% 
  group_by(PLAYERID,name) %>% 
   tally() %>%
  select(-n) %>% 
  collect()
) 
# user  system elapsed 
#    0.08    0.00    1.56

## error
system.time(
sl <-tbl(con, "playerGame") %>% 
  group_by(PLAYERID,name) %>% 
   slice(1) %>% 
  collect()
) 

## also refuses to work
system.time(
sl <-tbl(con, "playerGame") %>% 
  group_by(PLAYERID,name) %>% 
   slice(1) %>% 
  collapse()
) 

system.time(
di <-tbl(con, "playerGame") %>% 
  select(PLAYERID,name) %>% 
   distinct() %>% 
  collect()
) 



```

# works smoothly and fast until you have to turn into a dayt.frame

Lets look at an example from the site e.g career with list specified already

```{r}

## in shiny app replaced by interactive select
thePlayer <- "ROONEYX"

dfChart <- playerGame %>% 
    filter(PLAYERID==thePlayer) %>% 
    select(date=gameDate,Opponents,on,off=offA,Goals=Gls,Assists,Team=TEAMNAME,mins,plGameOrder,PLAYERID) %>% 
    mutate(points=Goals+Assists)

is.data.frame(fortify(dfChart))


dfChart %>% 
  fortify() %>% ## fortifying here is pretty much instantaneous
    plot_ly() %>% 
    add_markers(x=~date,y=~mins,color=~Team,
            hoverinfo = "text",
            text = ~paste(date,"<br>v ",Opponents,"<br>on:",on,"<br>off:",off,"<br>Goals:",Goals,"<br>Assists:",Assists),
            marker=list(size=~points*2+6)) %>% 
    layout(hovermode = "closest",
           title="",
           xaxis=list(title="Match day Squad Game order"),
           yaxis=list(title="Minutes played"
           )
    ) %>% 
    config(displayModeBar = F,showLink = F)

```

 works
 
 Now with the con directly in code
 
```{r}

dfTeamYear <- tbl(con, "summary") %>%
    filter(PLAYERID==thePlayer) %>% 
  fortify() %>% 
    mutate(apps=St+On,Gls=StGls+subGls,Pens=startPens+subPens,Points=Gls+Assists,ppg=round(Points*90/mins,2),janAge=as.integer(str_sub(season,1,4)),byear=as.integer(str_sub(born,1,4)),age=janAge-byear) %>%
    select(Season=season,Age=age,Team = TEAMNAME,Apps=apps,St,On,Off,Bench,Mins=mins,Gls,Assists,Points,ppg,Pens,Y,R,OG,MP) %>% 
    #ungroup() # looks like this was causing issue
    arrange(desc(Season))


 
  dfTeam <- dfTeamYear %>% 
    group_by(Team) %>%
    summarise(Apps=sum(Apps),St=sum(St),On=sum(On),Off=sum(Off),Bench=sum(Bench),Mins=sum(Mins),Goals=sum(Gls),
              Pens=sum(Pens),Assists=sum(Assists),Points=sum(Points),Y=sum(Y),R=sum(R),OG=sum(OG),MP=sum(MP))
  
 # #print(names(dfTeam))
  
  dfCareer <- dfTeamYear %>% 
   
    summarise(Apps=sum(Apps),St=sum(St),On=sum(On),Off=sum(Off),Bench=sum(Bench),Mins=sum(Mins),Goals=sum(Gls),
              Pens=sum(Pens),Assists=sum(Assists),Points=sum(Points),Y=sum(Y),R=sum(R),OG=sum(OG),MP=sum(MP))
  dfCareer$Team <- "Career"
  
  
  
  dfTeamYear  %>%  
    select(-MP) %>% 
    DT::datatable(class='compact stripe hover row-border order-column',rownames=FALSE,options= list(paging = FALSE, searching = FALSE,info=FALSE))
    


```
 
 Couple of other tbales not done here
 
```{r}
 df <-playerGame %>% 
    filter(PLAYERID==thePlayer) %>% 
  fortify() %>% 
    group_by(season,PLAYERID,name) %>% 
    select(Gls,Assists,mins) %>% 
    summarise(Goals=sum(Gls),Assists=sum(Assists),Points=Goals+Assists,Mins=sum(mins))%>% 
   # filter(Points!=0) %>% otherwise does not show zero values on year
    mutate(Gpm=90*Goals/Mins,Apm=90*Assists/Mins,Ppm=90*Points/Mins) %>% 
    ungroup() 
  
 #write_csv(df,"problem.csv")
#   
#   df %>% 
#     tauchart() %>% 
#     tau_point("season","Ppm", size="2") %>% 
#     # tau_line("season","Ppm") %>% 
#     tau_tooltip(c("Goals","Assists")) %>% 
#     tau_guide_x(label="") %>% 
#     tau_guide_y(label ='Points per 90 mins')
  df %>% 
  plot_ly() %>% 
  add_markers(x = ~season, y = ~Ppm,  hoverinfo = "text",
          marker=list(size=~Mins/10, sizemode="area"),
          text = ~paste(
            "<br>Goals: ",Goals,
            "<br>Assists: ",Assists,
            "<br>Points: ",Points,
            "<br>Minutes: ",Mins
          )) %>%
    layout(hovermode = "closest",
           title="Points per 90 mins by Season",
           xaxis=list(title=""),
           yaxis=list(title="Points per 90 mins",rangemode="tozero"
           )
    ) %>%  config(displayModeBar = F,showLink = F)
    
  
```
 
 
 ## Looks good
 
 